### 2019-07-31 honours-report/dependency.tex
thesis

% \begin{figure}[htbp] % here, top, bottom, page
\newcommand{\holtm}[1]{#1}
\newcommand{\holthm}[1]{#1}

\begin{tikzpicture}[>=stealth,auto, node distance=2cm]
\tikzstyle{block} = [draw, rectangle, minimum height=3em, minimum width=3em]
\tikzstyle{virtual} = [coordinate]
\tikzstyle{note} =[draw, rectangle, text width=6em,
                   inner sep=5pt, minimum height=1em, minimum width=1em]
\tikzstyle{element}=[circle, draw,, minimum size=12pt, inner sep=0pt]
    % Top Level
    \node [virtual] (ch1) at (4,18) {}; % maximum height = 18
    \node [note, text width=10em,  left=0.2 of ch1] (ch1al)
          {\texttt{IBC_DNF_EXISTS}};
    \node [note, text width=13em, right=0.2 of ch1] (ch1ar)
          {\HOLConst{Formal implementation}: \holtm{aksM n}};
    % First Level
    \node [virtual] (ch2) at (0,16) {};
    \node [virtual] (ch4) at (4.8,16) {};
    \node [virtual] (ch6) at (10,16) {};

    \node [virtual, below=0.3 of ch2]  (ch2a)    {};
    \node [note, text width=4em, left=0.1 of ch2a] (ch2al)  {\HOLConst{Monoids}};
    \node [note, text width=3em, right=0.1 of ch2a] (ch2ar) {\HOLConst{Groups}};
    \node [virtual, below=1.0 of ch2a] (ch2b)  {};
    \node [note, text width=3em, left=0.1 of ch2b] (ch2bl)  {\HOLConst{Rings}};
    \node [note, text width=3em, right=0.1 of ch2b] (ch2br) {\HOLConst{Fields}};
    \node [note, below=0.9 of ch2b] (ch2c) {\HOLConst{Polynomials}};
    \node [note, below=0.5 of ch2c] (ch2d) {\HOLConst{Quotient Rings}};
    \node [note, below=0.5 of ch2d] (ch2e) {\HOLConst{Number Theory}};

    \node [note, below=0 of ch4]  (ch4a)  {\HOLConst{Finite Fields}};
    \node [virtual, below=0.8 of ch4a] (ch4b)  {};
    \node [note, text width=4em, left=0.1 of ch4b] (ch4bl)  {\HOLConst{Subfields}};
    \node [note, right=0.1 of ch4b] (ch4br) {\HOLConst{Vector Spaces}};
    \node [virtual, below=0.9 of ch4b] (ch4c)  {};
    \node [note, text width=4em, left=0.1 of ch4c] (ch4cl)  {\HOLConst{Irreducibles}};
    \node [note, right=0.1 of ch4c] (ch4cr) {\HOLConst{Minimal Poly.}};
    \node [virtual, below=1.0 of ch4c] (ch4d)  {};
    \node [note, text width=4em, left=0.1 of ch4d] (ch4dl)  {\HOLConst{Existence}};
    \node [note, right=0.1 of ch4d] (ch4dr) {\HOLConst{Uniqueness}};
	\node [note, text width=8em, below=0.9 of ch4d] (ch4e)  {\HOLConst{Cyclotomic Factors}};

    \node [note, below=0 of ch6]  (ch6a) {\HOLConst{Machine} \HOLConst{Model}};
    \node [note, below=1 of ch6a] (ch6b) {\HOLConst{Recurrence} \HOLConst{Loops}};
    \node [note, below=1 of ch6b] (ch6c) {\HOLConst{Example:} \holtm{ulogM n}};

    % Second Level
    \node [virtual] (ch3) at (0,8.5) {};
    \node [virtual] (ch5) at (4.8,8.5) {};
    \node [virtual] (ch7) at (9.5,8.5) {};

    \node [note, below=0 of ch3]  (ch3a)   {\holtm{power_free n}};
    \node [note, below=0.5 of ch3a] (ch3b) {\holtm{aks_param n}};
    \node [note, text width=10em, below=0.5 of ch3b] (ch3c) {\holtm{poly_intro_range (ZN n) k n s}};
%    \node [note, text width=8em, below=0.5 of ch3c] (ch3d)  {\holtm{prime n ==> aks n}};
    \node [note, text width=8em, below=0.5 of ch3c] (ch3d)  {\holthm{AKSclean.aks_thm_1}};
    \node [note, text width=10em, below=0.5 of ch3d] (ch3e) {\holtm{poly_intro_range (ZN p) k n s}};
%    \node [note, text width=8em, below=0.5 of ch3e] (ch3f)  {\holtm{aks n ==> prime n}};
    \node [note, text width=8em, below=0.5 of ch3e] (ch3f)  {\holthm{AKSclean.aks_thm_2}};
    \node [note, text width=0, draw=none, left=0 of ch3e] (ch3x) {}; % dummy for chapter round

    \node [note, text width=5em, below=0 of ch5]  (ch5a)
           {\HOLConst{Introspective Relation:} \text{\ \; \holtm{n intro (X + |c|)}}};
    \node [virtual, below=1 of ch5a] (ch5b)  {};
    \node [note, text width=2em, left=0.1 of ch5b] (ch5bl)  {\holtm{NS}};
    \node [note, text width=2em, right=0.1 of ch5b] (ch5br) {\holtm{PS}};
    \node [virtual, below=1.5 of ch5b] (ch5c)  {};
    \node [note, text width=2em, left=0.1 of ch5c] (ch5cl)  {\holtm{MS}};
    \node [note, text width=2em, right=0.1 of ch5c] (ch5cr) {\holtm[polyh/h]{Q h}};
    \node [note, text width=2em, below=1 of ch5c] (ch5d)    {\HOLConst{Proof}};
    \node [note, text width=0, draw=none, right=0.3 of ch5d] (ch5x) {}; % dummy for chapter round

    \node [note, below=0 of ch7]  (ch7a)  {\holtm{power_freeM n}};
    \node [note, below=1 of ch7a] (ch7b)  {\holtm{paramM n}};
    \node [note, text width=9em, below=1 of ch7b] (ch7c)  {\holtm{poly_intro_rangeM n k s}};
    \node [note, text width=7em, below=1 of ch7c] (ch7d)  {\holtm{aksM n}: \HOLConst{analysis}};
    \node [note, text width=0, draw=none, left=0 of ch7c] (ch7x) {}; % dummy for chapter round

    % Final Level
    \node [virtual] (ch8) at (4,0) {};
%    \node [note, text width=8em, left=0.3 of ch8] (ch8a) {\holtm{aks n = prime n}};
%    \node [note, text width=12em, right=0.3 of ch8] (ch8b)
%          {\holtm{valueOf (aksM n) <=> aks n} \holtm{stepsOf o aksM IN big_O (\n. ulog n ** 21)}};
    \node [note, text width=10em, left=0.3 of ch8] (ch8a) {\holthm{AKSclean.aks_thm}};
    \node [note, text width=15em, right=0.3 of ch8] (ch8b)
          {\begin{HOLarray}\holthm{AKSclean.aksM_value_alt}\\
           \holthm{countAKS.aksM_steps_big_O}\end{HOLarray}};

   % Round up Chapters
    \node[draw, thick, green!60!black, rounded corners,
          label={\HOLConst{Ch~\ref{cha:introduction}: Introduction}},
          inner xsep=1em, inner ysep=1em, fit=(ch1al) (ch1ar)] {};
    \node[draw, thick, green!60!black, rounded corners,
          label={\HOLConst{Ch~\ref{cha:basic-algebra}: Basic Algebra}},
          inner xsep=1em, inner ysep=1em, fit=(ch2al) (ch2ar) (ch2e)] {};
    \node[draw, thick, green!60!black, rounded corners,
          label={\HOLConst{Ch~\ref{cha:aks-algorithm}: AKS Algorithm}},
          inner xsep=1em, inner ysep=1em, fit=(ch3a) (ch3c) (ch3f) (ch3x)] {};
    \node[draw, thick, green!60!black, rounded corners,
          label={\HOLConst{Ch~\ref{cha:advanced-algebra}: Advanced Algebra}},
          inner xsep=1em, inner ysep=1em, fit=(ch4a) (ch4dl) (ch4dr) (ch4e)] {};
    \node[draw, thick, green!60!black, rounded corners,
          label={\HOLConst{Ch~\ref{cha:aks-main-theorem}: AKS Main Theorem}},
          inner xsep=1em, inner ysep=1em, fit=(ch5a) (ch5d) (ch5x)] {};
    \node[draw, thick, green!60!black, rounded corners,
          label={\HOLConst{Ch~\ref{cha:complexity-models}: Complexity Models}},
          inner xsep=1em, inner ysep=1em, fit=(ch6a) (ch6c)] {};
    \node[draw, thick, green!60!black, rounded corners,
          label={\HOLConst{Ch~\ref{cha:aks-complexity}: AKS Complexity}},
          inner xsep=1em, inner ysep=1em, fit=(ch7a) (ch7c) (ch7d) (ch7x)] {};
    \node[draw, thick, green!60!black, rounded corners,
          label={\HOLConst{Ch~\ref{cha:conclusion}: Conclusion}},
          inner xsep=1em, inner ysep=1em, fit=(ch8a) (ch8b)] {};

   % Connections
    % (aks n) -> (aksM n)  
    \draw [->,purple] (ch1al) -- (ch1ar);
    % (aks n) -> Ch 3
    \node[virtual,below=0.6 of ch1al.south] (x1) {};
    \node[virtual,left=4.3 of x1] (x2) {};
    \node[virtual,below=7.5 of x2] (x3) {};
    \node[virtual,right=1.3 of x3] (x4) {};
    \node[virtual,below=0.2 of x4] (x5) {};
    \draw [->,purple,rounded corners] (ch1al.south) -- (x1) |- (x2) |- (x3) |- (x4) |- (x5);
    % (aksM n) -> Ch 7
    \node[virtual,below=0.6 of ch1ar.south] (x1) {};
    \node[virtual,right=5 of x1] (x2) {};
    \node[virtual,below=7.5 of x2] (x3) {};
    \node[virtual,left=1 of x3] (x4) {};
    \node[virtual,below=0.2 of x4] (x5) {};
    \draw [->,purple,rounded corners] (ch1ar.south) -- (x1) |- (x2) |- (x3) |- (x4) |- (x5);
    % (Proof) -> (aks n ==> prime n)
    \node[virtual,below=0.2 of ch5d.south] (x1) {};
    \draw [->,purple,rounded corners] (ch5d.south) -- (x1) |- (ch3f.east);
    % (prime n => aks n) -> (prime n = aks n)
    % (aks n => prime n) -> (prime n = aks n)
    \node[virtual,left=0.6 of ch3d.west] (x1) {};
    \node[virtual,above=0.8 of ch8a.north] (x2) {};
    \draw [->,purple,rounded corners] (ch3d.west) -- (x1) |- (ch8a.west);
    \draw [->,purple,rounded corners] (ch3f.south) |- (x2) -- (ch8a.north);
    % (power_freeM n) + (paramM n) + (introM ...) -> (aksM n: analysis)
    \node[virtual,right=0.75 of ch7a.east] (x1) {};
    \node[virtual,right=0.75 of ch7b.east] (x2) {};
    \node[virtual,right=0.2 of ch7c.east] (x3) {};
    \draw [->,purple,rounded corners] (ch7a.east) -- (x1) |- (ch7d.east);
    \draw [->,purple,rounded corners] (ch7b.east) -- (x2);
    \draw [->,purple,rounded corners] (ch7c.east) -- (x3);
    % (aksM n : analysis) -> (results)
    \node[virtual, above=0.5 of ch8b.north] (x1) {};
    \draw [->,purple,rounded corners] (ch7d.south) |- (x1) -- (ch8b.north);
    % (aks_param n) -> k -> (intro_range .. k)
    \node[virtual, right=0.5 of ch3b.east] (x1) {};
    \node[virtual, below=0.8 of x1] (x2) {};
    \node[virtual, right=0.4 of ch3b.east] (x3) {};
    \node[note,draw=none, text width=1em, above=0 of x3] () {$k$};
    \draw [->,purple,rounded corners] (ch3b.east) -- (x1) |- (x2);  
    % (Numbers) -> p + (intro ZN) -> (intro Zp)
    \node[virtual, below=0.2 of ch2e.300] (x1) {};
    \node[virtual, right=2.5 of x1] (x2) {};
    \node[virtual, right=0.6 of ch3c.east] (x3) {};
    \node[virtual, right=2.8 of x1] (x4) {};
    \node[note,draw=none, text width=1em, below=0 of x4] () {$p$};
    \draw [->,purple,rounded corners] (ch2e.300) -- (x1) |- (x2) |- (ch3e.east);
    \draw [->,purple,rounded corners] (ch3c.east) -- (x3);
    % (Numbers) -> (prime n => aks n) + FLT
    \node[virtual, below=0.3 of ch2e.250] (x1) {};
    \node[virtual, right=2.4 of x1] (x2) {};
    \node[virtual, above=0.3 of ch3d.north] (x3) {};
    \node[virtual, right=2.0 of x1] (x4) {};
    \node[note,draw=none, text width=1em, below=0 of x4] () {\scriptsize{\textsf{FLT}}};
    \draw [->,purple,rounded corners] (ch2e.250) -- (x1) |- (x2) |- (x3) |- (ch3d.north);
    % (Numbers) -> (aks_param n): existence of k
    \node[virtual, below=0.9 of ch2e.220] (x1) {};
    \node[virtual, right=2.3 of x1] (x2) {};
    \node[virtual, above=0.3 of ch3b.north] (x3) {};
    \node[virtual, right=0 of x1] (x4) {};
    \node[note, draw=none, text width=9em, above=0 of x4] () {\HOLConst{LCM bound}};
    \draw [->,purple,rounded corners] (ch2e.220) -- (x1) |- (x2) |- (x3) |- (ch3b.north);
    % (aks_param n) -> k -> (Mk)
    \node[virtual, right=1.5 of ch3b.east] (x1) {};
    \draw [->,purple,rounded corners] (ch3b.east) -- (x1) |- (ch5cl.west);
    % (Cyclo) -> h -> (Qh)
    \node[virtual, below=0.8 of ch4e.south] (x1) {};
    \node[virtual, right=1.85 of x1] (x2) {};
    \node[virtual, right=2.1 of x1] (x3) {};
    \node[note,draw=none, text width=1em, below=0 of x3] () {\holtm{polyh}};
    \draw [->,purple,rounded corners] (ch4e.south) -- (x1) |- (x2) |- (ch5cr.east);
    % (fields) -> (finite fields) + (subfields)
    \node[virtual, right=0.6 of ch2br.east] (x1) {};
    \node[virtual, left=1.2 of ch4bl.west] (x2) {};
    \draw [->,purple,rounded corners] (ch2br.east) -- (x1) |- (ch4a.west);
    \draw [->,purple,rounded corners] (x2) -- (ch4bl.west);
    % (Poly) -> (Irreducibles)
    \node[virtual, right=0.8 of ch2c.east] (x1) {};
    \draw [->,purple,rounded corners] (ch2c.east) -- (x1) |- (ch4cl.west);
    % (Quotient) -> (Existence)
    \node[virtual, right=0.8 of ch2d.east] (x1) {};
    \draw [->,purple,rounded corners] (ch2d.east) -- (x1) |- (ch4dl.west);
    % (power_free n) -> (power_freeM n)
    \node[element,left=1.0 of ch3a.west] (x1) {1};
    \node[element,left=0.8 of ch7a.west] (x2) {1};
    \draw [->,purple] (ch3a.west) -- (x1);
    \draw [->,purple] (x2) -- (ch7a.west);
    % (aks_param n) -> (paramM n)
    \node[element,left=1.0 of ch3b.west] (x1) {2};
    \node[element,left=0.8 of ch7b.west] (x2) {2};
    \draw [->,purple] (ch3b.west) -- (x1);
    \draw [->,purple] (x2) -- (ch7b.west);
    % (intro_range n ...) -> (intro_rangeM n ...)
    \node[element,left=0.2 of ch3c.west] (x1) {3};
    \node[element,left=0.2 of ch7c.west] (x2) {3};
    \draw [->,purple] (ch3c.west) -- (x1);
    \draw [->,purple] (x2) -- (ch7c.west);
    % (Model) + (Recurrence) -> (Example)
    \node[virtual, left=0.2 of ch6a.west] (x1) {};
    \node[virtual, left=0.2 of ch6b.west] (x2) {};
    \draw [->,purple] (ch6b.west) -- (x2);
    \draw [->,purple,rounded corners] (ch6a.west) -- (x1) |- (ch6c.west);
    % (Model) + (Recurrence) + (Example) -> Ch7
    \node[virtual, right=0.55 of ch6a.east] (x1) {};
    \node[virtual, right=0.55 of ch6b.east] (x2) {};
    \node[virtual, right=0.55 of ch6c.east] (x3) {};
    \draw [->,purple] (ch6a.east) -- (x1);
    \draw [->,purple] (ch6b.east) -- (x2);
    \draw [->,purple] (ch6c.east) -- (x3);
    % (Monoid) -> (Ring)
    \node[virtual, right=0.2 of ch2al.south] (x1) {};
    \draw [->,purple] (x1) -- (ch2bl.north);
    % (Group) -> (Field)
    \draw [->,purple] (ch2ar.south) -- (ch2br.north);
    % (Group) -> (Number)
    \node[virtual, below=0.15 of ch2ar.250] (x1) {};
    \node[virtual, left=2.8 of x1] (x2) {};
    \draw [->,purple,rounded corners] (ch2ar.250) -- (x1) |- (x2) |- (ch2e.west);
    % (Exist) + (Unique) -> (Cyclo)
    \node[virtual, above=0.4 of ch4e.north] (x1) {};
    \draw [->,purple,rounded corners] (ch4dl.south) |- (x1) -- (ch4e.north);
    \draw [->,purple,rounded corners] (ch4dr.south) |- (x1) -- (ch4e.north);
    % (Minimal) -> (Cyclo)
    \node[virtual, right=0.2 of ch4cr.east] (x1) {};
    \draw [->,purple,rounded corners] (ch4cr.east) -- (x1) |- (ch4e.east);
    % (aks n => prime n) -> (intro Zp) 
    \draw [->,purple] (ch3f.north) -- (ch3e.south);
    % (intro Zp) -> (Introspective)
    \node[virtual, below=0.2 of ch3e.300] (x1) {};
    \node[virtual, right=2.9 of x1] (x2) {};
    \draw [->,purple,rounded corners] (ch3e.300) -- (x1) |- (x2) |- (ch5a.west);
    % (Introspective) -> (N) + (P)
    \node[virtual, above=0.3 of ch5bl.north] (x1) {};
    \node[virtual, above=0.3 of ch5br.north] (x2) {};
    \draw [->,purple,rounded corners] (ch5a.south) |- (x1) -- (ch5bl.north);
    \draw [->,purple,rounded corners] (ch5a.south) |- (x2) -- (ch5br.north);
    % (N) -> (Mk), (P) -> (Qh)  
    \node[virtual, below=0.9 of ch5bl.south] (x1) {};
    \node[virtual, below=0.9 of ch5br.south] (x2) {};
    \draw [->,purple,rounded corners] (ch5bl.south) |- (x1);
    \draw [->,purple,rounded corners] (ch5br.south) |- (x2);
    % (Mk) + (Qh) -> (Proof)
    \node[virtual, above=0.4 of ch5d.north] (x1) {};
    \draw [->,purple,rounded corners] (ch5cl.south) |- (x1) -- (ch5d.north);
    \draw [->,purple,rounded corners] (ch5cr.south) |- (x1) -- (ch5d.north);
    % (Ring) + (Field) -> (Poly)
    \node[virtual, above=0.4 of ch2c.north] (x1) {};
    \draw [->,purple,rounded corners] (ch2bl.south) |- (x1) -- (ch2c.north);
    \draw [->,purple,rounded corners] (ch2br.south) |- (x1) -- (ch2c.north);
    % (Monoid) -> (Group)
    \draw [->,purple,rounded corners] (ch2al.east) -- (ch2ar.west);
    % (Ring) -> (Field)
    \draw [->,purple,rounded corners] (ch2bl.east) -- (ch2br.west);
    % (Poly) -> (Quotient)
    \draw [->,purple,rounded corners] (ch2c.south) |- (ch2d.north);
    % (FField) -> (Subfield)
    \node[virtual, above=0.2 of ch4bl.north] (x1) {};
    \draw [->,purple,rounded corners] (ch4a.south) |- (x1) |- (ch4bl.north);
    % (Vspace) -> (Subfield)
    \draw [->,purple,rounded corners] (ch4br.west) -- (ch4bl.east);
    % (Vspace) -> (Minimal)
    \draw [->,purple,rounded corners] (ch4br.south) -- (ch4cr.north);
    % (Irreducible) -> (Minimal)
    \draw [->,purple,rounded corners] (ch4cl.east) -- (ch4cr.west);
    % (Subfield) -> (Irreducible)
    \draw [->,purple,rounded corners] (ch4bl.south) -- (ch4cl.north);
    % (Irreducibles) -> (Exist)
    \draw [->,purple,rounded corners] (ch4cl.south) -- (ch4dl.north);
    % (Minimal) -> (Unique)
    \draw [->,purple,rounded corners] (ch4cr.south) -- (ch4dr.north);
    % PHP -> (Proof)
    \node[element, draw=none, right=0.4 of ch5d.east] (x1) {\scriptsize{\textsf{PHP}}};
    \draw [->,purple] (x1) -- (ch5d.east);
\end{tikzpicture}
% \caption{Dependency Diagram}
% \label{fig:dependency-diagram}
% \end{figure}

### 2019-08-23 honours-report/PhD proposal.tex
Add files for PhD proposal (incl. bib file)

\documentclass[11pt]{article}
\usepackage{hyperref}
\usepackage[top=1in, bottom=1in, left=.75in, right=1in]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy} \lhead{Proposal} \chead{PhD research proposal} \rhead{Yiming Xu u5943321}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{float}
\usepackage{amssymb}
\usepackage{extarrows}
\usepackage{enumerate}
\usepackage{amsthm}
\usepackage{polynom}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\usepackage{cleveref}
\usepackage{hyperref}
\usepackage{tikz-cd}
\usetikzlibrary{matrix, calc, arrows}
\usepackage{stmaryrd}
\usepackage[all]{xy}

\newenvironment{solution}
{\begin{proof}[Solution]}
	{\end{proof}}
\renewcommand{\thesubsection}{\thesection(\alph{subsection})}
\def\quotient#1#2{\raise1ex\hbox{$#1$}\Big/\lower1ex\hbox{$#2$}}
\setlength{\parskip}{1ex}
\setlength\parindent{0pt}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{remark}{Remark}[section]

\newcommand{\toby}[1]{\stackrel{#1}{\longrightarrow}}

\title{PhD research proposal}
\date{July, 2019}
\author{Yiming Xu}

\begin{document}
\maketitle
We will work in the area known as mechanised mathematics, aiming to bring computer-assisted rigour to mathematical proof. In particular, we will mechanise a significant result in the intersection of model theory and set theory. Model theory is the general study of the mathematical objects that satisfy sets of first order logic formulas (themselves called ``theories''). The result we propose to capture is $\mathfrak p= \mathfrak t$, where $\mathfrak p$ is the pseudo-intersection number and $\mathfrak t$ is the tower number. Here, the precise definitions of these two cardinals are given as:

\begin{itemize}
    \item A set  $A$ is \emph{almost contained} in another set $B$ if all but finitely members of $A$ are members of $B$. 
    \item A set $\mathcal T$ of sets where `almost contain' is a linear order on $\mathcal T$ is called a \emph{tower}.
    \item For a family $\{A_i\}_{i\in I}$ of sets, if $B$ is a set which is almost contained in every $A_i$, then $B$ is called a \emph{pseudo-intersection} of $\{A_i\}_{i\in I}$.
    \item For a family $\{A_i\}_{i\in I}$ of sets, if for every finite $\{A_n\}_{n\in\mathbb N}\subseteq \{A_i\}_{i\in I}$, we have $\bigcap \{A_n\}_{n\in\mathbb N}$ is infinite, then the family $\{A_i\}_{i\in I}$ is called has that \emph{finite intersection property}.
    \item $\mathfrak p$ is the cardinality of the smallest family which has the finite intersection property but which does not have a pseudo-intersection.
    \item $\mathfrak t$ is the cardinality of the smallest family which is a tower an does not have a pseudo-intersection.
\end{itemize}

Whether or not $\mathfrak p = \mathfrak t$ is the oldest problem on cardinal invariants of the continuum, which remained unsolved for more than 50 years, and has just been solved by Malliaris and Shelah in 2013.

With their proof, \textit{``the authors uncovered a novel connection between a model theoretic complexity hierarchy, the Keisler-order, and the theory of cardinal characteristics''}~\cite{Soukup2019}. In particular, they connect two rather isolated fields of mathematical logic: set theory and model theory. This result is thus a starting point of putting these two areas together, and by further exploring this connection, the community has great chance to prove more and more interesting results.

As this proof is so important, we want to make sure that it does not contain any subtle mistakes or hidden assumptions. Also, we want to make sure that we have understood it in every detail, and to record all those details in a machine. We propose to carry out this mechanisation in simple type theory, as implemented in the HOL4 proof assistant.

% The theorem prover HOL4 has some limitations. For instance, we cannot quantify over types, so we cannot simutinously refer to all type universe. But the HOL4 has obvious advantages as well, as is easier to use than many other theorem provers.

We wish to mechanise the proof of $\mathfrak p=\mathfrak t$ in HOL4:

\begin{itemize}

\item To ensure that the proof is absolutely correct and so we really know $\mathfrak p=\mathfrak t$.

\item The proof is complicated and enlightening, and many people be may interested in following it. But the paper itself does not give enough detail for a learner. If we mechanise it in HOL4, then if someone wants to learn it, they can easily run our proof interactively to see every detail clearly.

\item There has been very little work on mechanised model theory to date. The process of mechanising the proof of $\mathfrak p=\mathfrak t$  will require the construction of a library of model theory. A library capable of supporting the proof of $\mathfrak p=\mathfrak t$  will also be a valuable resource for others in this area. It is well known that model theory can also be applied to obtain significant results in other fields, including but not limited to set theory, algebra and analysis. This means the model theory library we will develop will be useful not only for mechanisation of more model theory, but will also have a good chance to be helpful for mechanising results in other domains.

\item In addition to the headline $\mathfrak p=\mathfrak t$ result, other theorems proved along the way will include important results about Keisler's order, some classification theory (dividing lines between levels of stability), some known relationships between other infinite cardinals and some more set theory (including forcing).

\item The simple type theory implemented in HOL4 has some limitations. For instance, it does not support dependent type theory, so we cannot quantify over types and hence we are not allowed to simultaneously refer to all type universes. By doing this project, we can demonstrate that HOL4's limited ability to do set theory will still be capable of mechanising the proof.

\end{itemize}
\section{Procedure}
Our main result $\mathfrak p=\mathfrak t$ is based on model theory, so we need to establish some model theory foundations before beginning our main proof. The first step will be to mechanise part of the standard textbook \emph{Model Theory} by Chang and Keisler~\cite{ChangKeisler1990}. In particular, we will need to reach section 7.1 about stability theory. A significantly important result we will mechanise is the proof of Morley's theorem, which says if a theory has exactly one model (up to isomorphism) in one uncountable cardinal, then it has exactly one model in every uncountable cardinal.

With the basics in the book mechanised, the next step is Keisler's order. Keisler's order is an order on countable theories which ranks classes of theories by complexity. In order to do this, we will need to enrich our existing theories on ultrafilters and ultraproducts, as well as some special definitions given by Keisler (OK filters, good filters, etc), which will lead to the following two central results:
\begin{itemize}
    \item Keisler's order has maximal and minimal classes.
    \item $\mathit{SOP_2}$ implies maximal theories in Keisler's order.
\end{itemize}

At this point, we can start with the main proof. We will follow the original proof by Malliaris and Shelah~\cite{MalliarisShelah2016}.

% The main material [1] will be a refinement of the original proof by M. Malliaris, S. Shelah [2]. Since [1] will use many result from [2], we will get part of the original proof mechanised as well.

\bibliography{everything}
\bibliographystyle{plain}

\end{document}

### 2019-11-11 p-modalmodels/paper.tex
basics on latex for working on the paper

\documentclass{llncs}

\begin{document}

\title{Mechanised Modal Model Theory}

\author{Yiming Xu and Michael Norrish}

\begin{abstract}
\end{abstract}

\maketitle

\section{Introduction}[1.5p]

% stress that previous work tends to focus on proof theory, e.g. tableaux
% focus on model theory is novel

\paragraph{Contributions}
This paper presents the first mechanised proofs of a number of basic results from the first two chapters of Blackburn~\emph{et al}~\cite{Blackburn} (e.g., bounded morphisms, bisimulations and the finite model property \emph{via} selection), as well as
\begin{itemize}
\item the saturation of ultraproduct models;
\item modal equivalence as bisimilarity between ultrafilter extensions; and
\item a close approximation of van Benthem's Characterization Theorem.
\end{itemize}
We also discuss where HOL's simple type theory lets us down: some standard results (including the best possible statement of van Benthem's Characterization Theorem) seem impossible to prove in our setting.

\subsection{Related Work}

\section{Syntax, Semantics and the Standard Translation}[1.5p]

\section{Basic Results}[2p]

\section{Mechanizing Ultrafilters and Ultraproducts}[1p]

\section{Bisimilarity and Ultrafilter Extensions}[2p]

\section{Saturation of Ultraproduct Models}[3p]

\section{Van Benthem's Characterization Theorem}[3p]

\section{Conclusion}[<1p]

\bibliography{../honours-report/everything}
\bibliographystyle{splncs04}

\end{document}

### 2019-11-17 p-modalmodels/paper.tex
paper 1st draft before 2nd meeting

@@
 \documentclass{llncs}
-
+\usepackage{holtexbasic}
+\usepackage{stmaryrd}
+\newtheorem{thm}{Theorem}[chapter]
+\newtheorem{defn}{Definition}[chapter]
+\newtheorem{lm}[thm]{Lemma}
+\newtheorem{coro}[thm]{Corollary}
+\newtheorem{prop}[thm]{Proposition}
+\newenvironment{holmath}{\begin{displaymath}\begin{array}{l}}{\end{array}\end{displaymath}\ignorespacesafterend}
+\renewcommand{\HOLConst}[1]{{\textsf{\upshape\small #1}}}
+\renewcommand{\HOLinline}[1]{\ensuremath{#1}}
+\renewcommand{\HOLFieldName}[1]{\HOLConst{#1}}
+\renewcommand{\HOLTokenLeftrec}{\ensuremath{\langle\!\langle}}
+\renewcommand{\HOLTokenRightrec}{\ensuremath{\rangle\!\rangle}}
+\newcommand{\holthmenv}[1]{\begin{array}[t]{l}#1\end{array}}
+\renewcommand{\HOLKeyword}[1]{\texttt{\upshape #1}}
+\renewcommand{\HOLTokenBar}{\mbox{$|$}}
 \begin{document}
@@
 
-\begin{abstract}
-\end{abstract}
+%\begin{abstract}
+%\end{abstract}
 
@@
 
-\section{Introduction}[1.5p]
+\section{Introduction}[1p]
 
@@
 \section{Syntax, Semantics and the Standard Translation}[1.5p]
+In our formalization, we only consider the basic modal language, in which the only primitive modal operator is the `$\Diamond$'.
+For a type $\alpha$, an $\alpha$-modal formula is either of form \HOLinline{\HOLConst{VAR}\;\HOLFreeVar{p}}, where $p$ is of type $\alpha$, or a disjunction $\phi\lor \psi$ of two $\alpha$-modal formulas, or the falsity $\bot$, or a negation $\lnot \phi$ of an $\alpha$-modal formula $\phi$, or, finally, of the form $\Diamond \phi$ where $\phi$ is an $\alpha$-modal formula.
+We create a data type called `form' of the formulas of this modal language.
+\begin{defn}
+{\upshape\cite[Definition 1.9]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\alpha}\;\HOLTyOp{form}\;=\;\HOLConst{VAR}\;\ensuremath{\alpha}\;\HOLTokenBar{}\;\HOLConst{DISJ}\;(\ensuremath{\alpha}\;\HOLTyOp{form})\;(\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLTokenBar{}\;\HOLSymConst{\ensuremath{\bot}}\;\HOLTokenBar{}\;(\HOLSymConst{\HOLTokenNeg{}})\;(\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLTokenBar{}\;\HOLConst{\ensuremath{\Diamond}}\;(\ensuremath{\alpha}\;\HOLTyOp{form})
+\end{holmath}
+\end{defn}
 
-\section{Basic Results}[2p]
+A model where these formulas can be interpreted consists of a \emph{frame} and a \emph{valuation}, where \HOLinline{\HOLConst{frame}} is a set with a relation on it. 
+\begin{defn}
+{\upshape\cite[Definition 1.19]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\beta}\;\HOLTyOp{frame} = \HOLTokenLeftrec{}\;\HOLFieldName{world}\;{:}\;\ensuremath{\beta}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool};\;\HOLFieldName{rel}\;{:}\;\ensuremath{\beta}\;\ensuremath{\rightarrow}\;\ensuremath{\beta}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}\;\HOLTokenRightrec{}\\
+  (\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model} = \HOLTokenLeftrec{}\;\HOLFieldName{frame}\;{:}\;\ensuremath{\beta}\;\HOLTyOp{frame};\;\HOLFieldName{valt}\;{:}\;\ensuremath{\alpha}\;\ensuremath{\rightarrow}\;\ensuremath{\beta}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}\;\HOLTokenRightrec{}
+\end{holmath}
+\end{defn}
+In the rest of the paper, the field \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}}$.\sf valt$ of a model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} will be called the \emph{valuation}, and \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}}, \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{R}}}} and \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{V}}}} are used to denote the world set, the relation, and the valuation of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} respectively.
+When we say a $(\HOLinline{\ensuremath{\alpha}},\HOLinline{\ensuremath{\beta}})$-model, we mean a model for $\alpha$-formulas with a $\beta$-set as its underlying set.
+The interpretation of $\alpha$-modal formulas on an $(\HOLinline{\ensuremath{\alpha}},\HOLinline{\ensuremath{\beta}})$-model is given by the predicate \emph{satisfaction}.
+We read `\HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}}' as `$\phi$ is satisfied at the world $w$ in \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}}'.
+\begin{defn}
+{\upshape\cite[Definition 1.20]{Blackburn}}
+\begin{holmath}
+\begin{array}{rcl}
+  \HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLConst{VAR}\;\HOLFreeVar{p}\;&\HOLTokenDefEquality{}&\;\HOLFreeVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{V}}}\;\HOLFreeVar{p}\\
+\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLSymConst{\ensuremath{\bot}}\;&\HOLTokenDefEquality{}&\;\HOLFreeVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{F}\\
+\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLSymConst{\HOLTokenNeg{}}\HOLFreeVar{\ensuremath{\phi}}\;&\HOLTokenDefEquality{}&\;\HOLFreeVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\not\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\\
+\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi\sb{1}}}\HOLSymConst{\ensuremath{\lor}}\HOLFreeVar{\ensuremath{\phi\sb{2}}})\;&\HOLTokenDefEquality{}&\;\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{1}}}\;\HOLSymConst{\HOLTokenDisj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{2}}}\\
+\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi}}\;&\HOLTokenDefEquality{}&\;\HOLFreeVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{v}.\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{R}}}\;\HOLFreeVar{w}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{v}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}
+\end{array}
+\end{holmath}
+\end{defn}
+If \HOLinline{\HOLFreeVar{\ensuremath{\phi\sb{1}}}}, \HOLinline{\HOLFreeVar{\ensuremath{\phi\sb{2}}}} are \HOLinline{\HOLFreeVar{\ensuremath{\alpha}}}-formulas, we say they are \emph{equivalent} if for every $(\alpha,\beta)$-model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} and every world \HOLinline{\HOLFreeVar{w}} in it, we have \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{1}}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{2}}}}.
+\begin{defn}[Equivalence]
+\begin{holmath}
+  (\HOLFreeVar{\ensuremath{\phi\sb{1}}} :\ensuremath{\alpha}\;\HOLTyOp{form})\equiv\sb{({:}\ensuremath{\beta})}(\HOLFreeVar{\ensuremath{\phi\sb{2}}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLTokenDefEquality{}\\
+\;\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model})\;(\HOLBoundVar{w} :\ensuremath{\beta}).\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{1}}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{2}}}\\[3mm]
+\end{holmath}
+\end{defn}
+We cannot omit the type parameter \HOLinline{({:}\ensuremath{\beta})} in the definition, since then there will be a type, namely the type of the underlying set of the models we are talking about, that only appears on the right-hand side but not on the left-hand side of the definition, which is not allowed in HOL.
+Also, we are not allowed to quantify over types, so it is also impossible to define the equivalence to be \HOLinline{\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{\ensuremath{\mu}}.\;\HOLFreeVar{\ensuremath{\phi\sb{1}}}\equiv\sb{\HOLBoundVar{\ensuremath{\mu}}}\HOLFreeVar{\ensuremath{\phi\sb{2}}}}, where $\mu$ denotes a type.
+Therefore, this definition is not encoding the equivalence in mathematical sense precisely: when we mention equivalence of formulas in usual mathematical language, we are implicitly referring to the class of all models, but the constraint here bans us from talking about all models of all possible types at once.
 
-\section{Mechanizing Ultrafilters and Ultraproducts}[1p]
+A modal formula can be translated into a first-order formula via the \emph{standard translation}.
+A translated formula has at most one free variable.
+For a modal formula \HOLinline{\HOLFreeVar{\ensuremath{\phi}}}, \HOLinline{{\sf ST}\sb{\HOLFreeVar{x}}\;\HOLFreeVar{\ensuremath{\phi}}} is the standard translation of \HOLinline{\HOLFreeVar{\ensuremath{\phi}}} using \HOLinline{\HOLFreeVar{x}} as the only free variable.
+\begin{defn}
+{\upshape\cite[Definition 2.45 (Standard Translation)]{Blackburn}}
+\begin{holmath}
+\begin{array}{rcl}
+  {\sf ST}\sb{\HOLFreeVar{x}}\;(\HOLConst{VAR}\;\HOLFreeVar{p})\;&\HOLTokenDefEquality{}&\;\HOLConst{\ensuremath{{}\sp{\mathsf{f}}\mathsf{P}}}\;\HOLFreeVar{p}\;(\HOLConst{\ensuremath{{}^{\sf f}{\sf VAR}}}\;\HOLFreeVar{x})\\
+{\sf ST}\sb{\HOLFreeVar{x}}\;\HOLSymConst{\ensuremath{\bot}}\;&\HOLTokenDefEquality{}&\;\HOLConst{\ensuremath{{}^{\sf f}\bot}}\\
+{\sf ST}\sb{\HOLFreeVar{x}}\;(\HOLSymConst{\HOLTokenNeg{}}\HOLFreeVar{\ensuremath{\phi}})\;&\HOLTokenDefEquality{}&\;\HOLConst{\ensuremath{{^{\sf f}}{\lnot}}}\;({\sf ST}\sb{\HOLFreeVar{x}}\;\HOLFreeVar{\ensuremath{\phi}})\\
+{\sf ST}\sb{\HOLFreeVar{x}}\;(\HOLFreeVar{\ensuremath{\phi}}\HOLSymConst{\ensuremath{\lor}}\HOLFreeVar{\ensuremath{\psi}})\;&\HOLTokenDefEquality{}&\;{\sf ST}\sb{\HOLFreeVar{x}}\;\HOLFreeVar{\ensuremath{\phi}}\HOLSymConst{\ensuremath{\ {}^{\sf f} {\lor} \ }}{\sf ST}\sb{\HOLFreeVar{x}}\;\HOLFreeVar{\ensuremath{\psi}}\\
+{\sf ST}\sb{\HOLFreeVar{x}}\;(\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi}})\;&\HOLTokenDefEquality{}&\;\HOLConst{\ensuremath{ {}^{\sf f}{\exists}}}\;(\HOLFreeVar{x}\;\HOLSymConst{\ensuremath{+}}\;\HOLNumLit{1})\;(\HOLConst{\ensuremath{{}^{\sf f}{\sf R}}}\;(\HOLConst{\ensuremath{{}^{\sf f}{\sf VAR}}}\;\HOLFreeVar{x})\;(\HOLConst{\ensuremath{{}^{\sf f}{\sf VAR}}}\;(\HOLFreeVar{x}\;\HOLSymConst{\ensuremath{+}}\;\HOLNumLit{1}))\HOLSymConst{\ensuremath{\ {}^{\sf f}{\land} \ }}{\sf ST}\sb{\HOLFreeVar{x}\;\HOLSymConst{\ensuremath{+}}\;\HOLNumLit{1}}\;\HOLFreeVar{\ensuremath{\phi}})
+\end{array}
+\end{holmath}
+\end{defn}
+According to the semantic interpretation of the `\HOLinline{\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi}}}', we translate \HOLinline{\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi}}} into the existential formula.
+To make sure that we use a fresh variable symbol that is not the same as the variable $x$ which is marking the current state, we use $x+1$ as our new variable symbol.
+The standard translation gives a first-order reformulation of satisfaction of modal formulas:
+\begin{prop}\label{2.47}
+{\upshape\cite[Theorem 2.47 (i)]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLConst{mm2folm}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}},(\HOLTokenLambda{}\HOLBoundVar{n}.\;\HOLFreeVar{w})\ensuremath{\vDash}\;{\sf ST}\sb{\HOLFreeVar{x}}\;\HOLFreeVar{\ensuremath{\phi}}
+\end{holmath}
+\end{prop} 
+Here \HOLinline{\HOLConst{mm2folm}} is the function that turns a modal model into a first-order model, defined as:
+\begin{holmath}
+  \HOLConst{mm2folm}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLTokenDefEquality{}\\
+\;\;\HOLTokenLeftrec{}\HOLFieldName{Dom}\;:=\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}};\;\HOLFieldName{Fun}\;:=\;(\HOLTokenLambda{}\HOLBoundVar{n}\;\HOLBoundVar{l}.\;\HOLConst{CHOICE}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}});\\
+\;\;\HOLFieldName{Pred}\;:=\\
+\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{p}\;\HOLBoundVar{zs}.\\
+\;\;\;\;\;\;\;\;\;\HOLKeyword{case}\;\HOLBoundVar{zs}\;\HOLKeyword{of}\\
+\;\;\;\;\;\;\;\;\;\;\;[]\;\HOLTokenImp{}\;\HOLConst{F}\\
+\;\;\;\;\;\;\;\;\;\HOLTokenBar{}\;[\HOLBoundVar{w}\sb{\mathrm{1}}]\;\HOLTokenImp{}\;\HOLBoundVar{w}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{V}}}\;\HOLBoundVar{p}\;\HOLBoundVar{w}\sb{\mathrm{1}}\\
+\;\;\;\;\;\;\;\;\;\HOLTokenBar{}\;[\HOLBoundVar{w}\sb{\mathrm{1}};\;\HOLBoundVar{w}\sb{\mathrm{2}}]\;\HOLTokenImp{}\;\HOLBoundVar{p}\;\HOLSymConst{=}\;\HOLNumLit{0}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{R}}}\;\HOLBoundVar{w}\sb{\mathrm{1}}\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{w}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\\
+\;\;\;\;\;\;\;\;\;\HOLTokenBar{}\;\HOLBoundVar{w}\sb{\mathrm{1}}\HOLSymConst{::}\HOLBoundVar{w}\sb{\mathrm{2}}\HOLSymConst{::}\HOLBoundVar{w}\sb{\mathrm{3}}\HOLSymConst{::}\HOLBoundVar{w}\sb{\mathrm{4}}\;\HOLTokenImp{}\;\HOLConst{F})\HOLTokenRightrec{}
+\end{holmath}
 
-\section{Bisimilarity and Ultrafilter Extensions}[2p]
+\section{Basic Results}%[2p]
+We discuss some highlights of formalized results in section 2.1-2.3 in \cite{Blackburn} below.
+\subsection{Tree-like property}
+A tree-like model is a model whose underlying frame is a tree. If $H$ is a tree with root $r$, we write \HOLinline{\HOLConst{tree}\;\HOLFreeVar{H}\;\HOLFreeVar{r}} in HOL:
+\begin{defn}
+{\upshape\cite[Definition 1.7]{Blackburn}}
+\begin{holmath}
+   \HOLConst{tree}\;\HOLFreeVar{H}\;\HOLFreeVar{r}\;\HOLTokenDefEquality{}\\
+\;\;\HOLFreeVar{r}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{w}.\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenImp{}}\;{\HOLFreeVar{H}.\HOLFieldName{rel}}\ensuremath{\mid}\sb{\HOLFreeVar{H}.\HOLFieldName{world}}\HOLSymConst{\HOLTokenSupStar{}}\;\HOLFreeVar{r}\;\HOLBoundVar{w})\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{w}.\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenNeg{}}\HOLFreeVar{H}.\HOLFieldName{rel}\;\HOLBoundVar{w}\;\HOLFreeVar{r})\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{w}.\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLFreeVar{r}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenUnique{}}\HOLBoundVar{w}\sb{\mathrm{0}}.\;\HOLBoundVar{w}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{H}.\HOLFieldName{rel}\;\HOLBoundVar{w}\sb{\mathrm{0}}\;\HOLBoundVar{w}
+\end{holmath}
+\end{defn}
+The tree-like property says an satisfiable modal formula can be satisfied in a tree-like model. It is proved in HOL as:
+\begin{prop}
+{\upshape\cite[Proposition 2.15]{Blackburn}}\label{2.15}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model}),(\HOLFreeVar{w} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :(\ensuremath{\alpha},\;\ensuremath{\beta}\;\HOLTyOp{list})\;\HOLTyOp{model})\;(\HOLBoundVar{r} :\ensuremath{\beta}\;\HOLTyOp{list}).\;\HOLConst{tree}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{frame}\;\HOLBoundVar{r}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{r}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}
+\end{holmath}
+\end{prop}
+The world set of the tree like model constructed from \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} according to the standard proof of this theorem is a set of lists of worlds in \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}}, that is why passing to a tree-like model does not preserve the type.
 
-\section{Saturation of Ultraproduct Models}[3p]
+\subsection{Bisimulation}
+The definition of bisimulation in HOL is just a direct translation of its mathematical definition: 
+\begin{defn}
+{\upshape\cite[Definition 2.16 (Bisimulations)]{Blackburn}}
+\begin{holmath}
+  \HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\overset{\HOLFreeVar{Z}}\leftrightarroweq\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\;\HOLTokenDefEquality{}\\
+\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{w}\sb{\mathrm{1}}\;\HOLBoundVar{w}\sb{\mathrm{2}}.\\
+\;\;\;\;\;\;\HOLBoundVar{w}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{Z}\;\HOLBoundVar{w}\sb{\mathrm{1}}\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{p}.\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLBoundVar{w}\sb{\mathrm{1}}\ensuremath{\Vdash}\;\HOLConst{VAR}\;\HOLBoundVar{p}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLBoundVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLConst{VAR}\;\HOLBoundVar{p})\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{v}\sb{\mathrm{1}}.\\
+\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLBoundVar{v}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\HOLSymConst{\ensuremath{\sp{R}}}\;\HOLBoundVar{w}\sb{\mathrm{1}}\;\HOLBoundVar{v}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{v}\sb{\mathrm{2}}.\;\HOLBoundVar{v}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{Z}\;\HOLBoundVar{v}\sb{\mathrm{1}}\;\HOLBoundVar{v}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{R}}}\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLBoundVar{v}\sb{\mathrm{2}})\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{v}\sb{\mathrm{2}}.\\
+\;\;\;\;\;\;\;\;\;\;\;\;\HOLBoundVar{v}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{R}}}\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLBoundVar{v}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{v}\sb{\mathrm{1}}.\;\HOLBoundVar{v}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{Z}\;\HOLBoundVar{v}\sb{\mathrm{1}}\;\HOLBoundVar{v}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\HOLSymConst{\ensuremath{\sp{R}}}\;\HOLBoundVar{w}\sb{\mathrm{1}}\;\HOLBoundVar{v}\sb{\mathrm{1}}
+\end{holmath}
+\end{defn}
 
-\section{Van Benthem's Characterization Theorem}[3p]
+It is trivial to prove by induction that bisimilar worlds are modal equivalent.
+As the most significant theorem on the basic theory of bisimulations, we proved the Hennessy-Milner theorem as, which says modal equivalence and bisimulation on \emph{image finite} models are the same thing.
+An image-finite model is a model where every world can only be related to finitely many worlds.
+The Hennessy-Milner Theorem in HOL looks like:
+
+\begin{thm}
+{\upshape\cite[Theorem 2.24 (Hennessy-Milner Theorem)]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{image_finite}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{image_finite}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{w}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\sb{\mathrm{1}}\ensuremath{\leftrightsquigarrow}\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLFreeVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\HOLSymConst{,}\HOLFreeVar{w}\sb{\mathrm{1}}\HOLSymConst{\ensuremath{\leftrightarroweq}}\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{,}\HOLFreeVar{w}\sb{\mathrm{2}})
+\end{holmath}
+\end{thm}
+
+\subsection{Finite model property}
+There are two classical approach of constructing finite models using model theory, namely via selection and via filtration. 
+\subsubsection{Selection}
+Given \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\sb{\mathrm{1}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}}, where $\phi$ has degree $k$, we can construct \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}}, \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}}, \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}} and \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}} comsecutively, such that \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}} is the finite model we want, where:
+\begin{itemize}
+\item \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} is the tree-like model obtained from \ref{2.15} with root \HOLinline{\HOLFreeVar{w}\sb{\mathrm{2}}} such that \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLFreeVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}}.
+\item \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}} is the restriction of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} to height $k$.
+\item \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}} is obtained from \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}} by modifying the valuation into \HOLinline{\HOLTokenLambda{}\HOLBoundVar{p}\;\HOLBoundVar{v}.\;\HOLKeyword{if}\;\HOLBoundVar{p}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLConst{prop_letters}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLKeyword{then}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}\HOLSymConst{\ensuremath{\sp{V}}}\;\HOLBoundVar{p}\;\HOLBoundVar{v}\;\HOLKeyword{else}\;\HOLConst{F}}.
+\end{itemize}
+The construction of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}} requires a lemma: 
+\begin{lm}
+{\upshape\cite[Proposition 2.29]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{FINITE}\;(\HOLFreeVar{\ensuremath{\Phi}} :\ensuremath{\alpha}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{INFINITE}\;\ensuremath{{\cal{U}}}({:}\ensuremath{\beta})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{n} :\HOLTyOp{num}).\;\HOLConst{FINITE}\;{\HOLTokenLeftbrace{}\HOLBoundVar{\ensuremath{\phi}}\;\HOLTokenBar{}\;\HOLConst{DEG}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenLeq{}}\;\HOLBoundVar{n}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{prop_letters}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\Phi}}\HOLTokenRightbrace{}} /{\equiv\sb{(:\beta)}}
+\end{holmath}
+\end{lm}
+We require the assumption that the universe of $\beta$ is infinite since we relies on the fact that two modal formulas \HOLinline{\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{1}}}} and \HOLinline{\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{2}}}} are equivalent if and only if \HOLinline{\HOLFreeVar{\ensuremath{\phi\sb{1}}}} and \HOLinline{\HOLFreeVar{\ensuremath{\phi\sb{2}}}} are equivalent.
+This is easy to prove using set theory.
+However, in simple type theory, the proof of  \HOLinline{\HOLFreeVar{\ensuremath{\phi\sb{1}}}\equiv\sb{({:}\ensuremath{\beta})}\HOLFreeVar{\ensuremath{\phi\sb{2}}}} iff \HOLinline{\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{1}}}\equiv\sb{({:}\ensuremath{\beta})}\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{2}}}} requires the infiniteness of the type universe of \HOLinline{\ensuremath{\beta}}.
+
+By the lemma above, the set
+\HOLinline{\HOLFreeVar{\ensuremath{\Delta}}\;\HOLSymConst{=}\;\HOLTokenLeftbrace{}\HOLBoundVar{\ensuremath{\psi}}\;\HOLTokenBar{}\;\HOLConst{DEG}\;\HOLBoundVar{\ensuremath{\psi}}\;\HOLSymConst{\HOLTokenLeq{}}\;\HOLFreeVar{k}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{prop_letters}\;\HOLBoundVar{\ensuremath{\psi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\Phi}}\HOLTokenRightbrace{}}$/\equiv_{(:\HOLinline{\ensuremath{\beta}\;\HOLTyOp{list}})}$ is finite.
+Let \HOLinline{\HOLFreeVar{\ensuremath{\Gamma}}} be the set of modal formulas starting with a diamond and let $R$ denote \HOLinline{\HOLTokenLeftbrace{}\HOLConst{CHOICE}\;(\HOLBoundVar{A}\;\HOLSymConst{\HOLTokenInter{}}\;\HOLFreeVar{\ensuremath{\Gamma}})\;\HOLTokenBar{}\;\HOLBoundVar{A}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\Delta}}\HOLTokenRightbrace{}\;\HOLConst{\ensuremath{\setminus}}\;\HOLTokenLeftbrace{}\HOLSymConst{\HOLTokenEmpty{}}\HOLTokenRightbrace{}}, we define \HOLinline{\HOLFreeVar{\ensuremath{S\sp{\sf s}}}} to be the primitive recursion function:
+\begin{holmath}
+ \HOLConst{PRIM_REC}\;\HOLTokenLeftbrace{}\HOLFreeVar{w}\sb{\mathrm{2}}\HOLTokenRightbrace{}\\
+\;\;(\HOLTokenLambda{}\HOLBoundVar{s}\sb{\mathrm{0}}\;\HOLBoundVar{n}.\\
+\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLConst{CHOICE}\;\HOLBoundVar{us}\;\HOLTokenBar{}\\
+\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{\ensuremath{\phi}}\;\HOLBoundVar{v}.\\
+\;\;\;\;\;\;\;\;\;\;\;\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}},\HOLBoundVar{v}\ensuremath{\Vdash}\;\HOLSymConst{\ensuremath{\Diamond}}\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\ensuremath{\Diamond}}\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{R}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{s}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\;\;\;\;\;\;\HOLBoundVar{us}\;\HOLSymConst{=}\;\HOLTokenLeftbrace{}\HOLBoundVar{u}\;\HOLTokenBar{}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}\HOLSymConst{\ensuremath{\sp{R}}}\;\HOLBoundVar{v}\;\HOLBoundVar{u}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{u}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}},\HOLBoundVar{u}\ensuremath{\Vdash}\;\HOLBoundVar{\ensuremath{\phi}}\HOLTokenRightbrace{}\HOLTokenRightbrace{})
+\end{holmath}
+Then the world set of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}} is \HOLinline{\HOLConst{\ensuremath{\bigcup}}\;\HOLTokenLeftbrace{}\HOLFreeVar{\ensuremath{S\sp{\sf s}}}\;\HOLBoundVar{i}\;\HOLTokenBar{}\;\HOLBoundVar{i}\;\HOLSymConst{\HOLTokenLeq{}}\;\HOLFreeVar{k}\HOLTokenRightbrace{}}.
+We proved \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}},\HOLFreeVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}} by proving
+\begin{holmath}
+ \HOLTokenLambda{}\HOLBoundVar{n}\;\HOLBoundVar{a}\sb{\mathrm{1}}\;\HOLBoundVar{a}\sb{\mathrm{2}}.\\
+\;\;\;\;\HOLBoundVar{a}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{a}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{height}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}\;\HOLFreeVar{w}\sb{\mathrm{2}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}\;\HOLBoundVar{a}\sb{\mathrm{1}}\;\HOLSymConst{=}\;\HOLConst{height}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}\;\HOLFreeVar{w}\sb{\mathrm{2}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}\;\HOLBoundVar{a}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\HOLConst{height}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}\;\HOLFreeVar{w}\sb{\mathrm{2}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}\;\HOLBoundVar{a}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenLeq{}}\;\HOLFreeVar{k}\;\HOLSymConst{\ensuremath{-}}\;\HOLBoundVar{n}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{\ensuremath{\phi}}.\;\HOLConst{DEG}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenLeq{}}\;\HOLBoundVar{n}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{prop_letters}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{s}\;\HOLSymConst{\HOLTokenImp{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}},\HOLBoundVar{a}\sb{\mathrm{1}}\ensuremath{\Vdash}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}},\HOLBoundVar{a}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLBoundVar{\ensuremath{\phi}})
+\end{holmath}
+is an \emph{n-bisimulation} relating the two copies of \HOLinline{\HOLFreeVar{w}\sb{\mathrm{2}}} in \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}} and in \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}}.
+
+As we used \ref{2.15}, we change the type of the model by passing to a finite model via selection:
+\begin{thm}
+{\upshape\cite[Theorem 2.34 (Finite model property, via selection)]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model}),(\HOLFreeVar{w}\sb{\mathrm{1}} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :(\ensuremath{\alpha},\;\ensuremath{\beta}\;\HOLTyOp{list})\;\HOLTyOp{model})\;(\HOLBoundVar{v} :\ensuremath{\beta}\;\HOLTyOp{list}).\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{v}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}
+\end{holmath}
+\end{thm}
+
+\subsubsection{Filtration}
+Given \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}}, the finite model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} constructed by selection is the filtration of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}} by the set of subformulas of \HOLinline{\HOLFreeVar{\ensuremath{\phi}}}.
+In HOL, we write \HOLinline{\HOLConst{FLT}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{\ensuremath{\Sigma}}} for the filtration of the model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} via the set \HOLinline{\HOLFreeVar{\ensuremath{\Sigma}}}.
+The formalization is almost a direct translation of the mathematical proof in \cite{Blackburn}.
+For the only difference: the world set of the filtrated model using the standard proof is a set of equivalence classes of worlds.
+To make the type of the finite model same as the model we start with, instead of taking the set of equivalence classes as the world set, we pick only one representative from each equivalence class.
+Hence we can prove:
+\begin{thm}
+{\upshape\cite[Theorem 2.41 (Finite model property, via filtration)]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\sb{\mathrm{1}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\;\HOLBoundVar{w}\sb{\mathrm{2}}.\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLBoundVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}
+\end{holmath}
+\end{thm}
+where \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} above is taken as \HOLinline{\HOLConst{FLT}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\;(\HOLConst{subforms}\;\HOLFreeVar{\ensuremath{\phi}})}, and has the same type as \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}}.
+
+\section{Mechanizing Ultrafilters and Ultraproducts}[3p]
+Some results in section 2.5-2.7 relies on some theorems about ultrafilters and ultraproduct. 
+\subsection{ultrafilters}
+Given a non-empty set $J$, a set $L\subseteq {\mathcal P}(J)$   is a \emph{filter} if it contains $J$ itself, is closed under binary intersection, and is closed upward. 
+\begin{defn}
+{\upshape\cite[Definition A.12 (Filters)]{Blackburn}}
+\begin{holmath}
+  \HOLConst{filter}\;\HOLFreeVar{L}\;\HOLFreeVar{J}\;\HOLTokenDefEquality{}\\
+\;\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLSymConst{\HOLTokenEmpty{}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{L}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLConst{POW}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{L}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X}\;\HOLBoundVar{Y}.\;\HOLBoundVar{X}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{L}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{Y}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{L}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLBoundVar{X}\;\HOLSymConst{\HOLTokenInter{}}\;\HOLBoundVar{Y}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{L})\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X}\;\HOLBoundVar{Z}.\;\HOLBoundVar{X}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{L}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{X}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLBoundVar{Z}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{Z}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLBoundVar{Z}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{L}
+\end{holmath}
+\end{defn}
+We call $L$ a \emph{proper filter} if $L$ is not the whole power set. An \emph{ultrafilter} is a filter $U$ such that for every $X\subseteq J$, exactly one of $X$ or $J\setminus X$ is in $U$.
+
+The ultrafilter theorem, which says every proper filter is contained in an ultrafilter, is proved as follows:
+\begin{thm}
+{\upshape\cite[Fact A.14, first half]{Blackburn}}\label{A.14}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{proper_filter}\;\HOLFreeVar{L}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{U}.\;\HOLConst{ultrafilter}\;\HOLBoundVar{U}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{L}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLBoundVar{U}
+\end{holmath}
+\end{thm}
+A subset $A$ of the power set on $J$ has \emph{finite intersection property} if once we take the intersection of a finite, nonempty family in $A$, the resultant set is nonempty.
+\begin{defn}
+{\upshape\cite[Definition A.13 (Finite Intersection Property)]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{FIP}\;\HOLFreeVar{A}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenEquiv{}}\\
+\;\;\;\;\;\;\;\HOLFreeVar{A}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLConst{POW}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{B}.\;\HOLBoundVar{B}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{A}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{B}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{B}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLSymConst{\HOLTokenEmpty{}}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\ensuremath{\bigcap}}\;\HOLBoundVar{B}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLSymConst{\HOLTokenEmpty{}}
+\end{holmath}
+\end{defn}
+As a corollary of ultrafilter theorem, a set with finite intersection property is contained in an ultrafilter.
+
+\subsection{Ultraproducts}
+The notion of ultraproducts are defined for sets, modal models, and first-order models. 
+\subsubsection{Ultraproduct of sets}
+The ultraproduct of a family of sets $(A_j)_{j\in J}$, where $J$ is non-empty and each $A_j$ is non-empty, is defined as a quotient of the cartesian product of the family.
+The Cartesian product of the family  $(A_j)_{j\in J}$ is the set of functions $f$ with domain $J$ such that for all $j\in J$, $f(j)\in A_j$.
+\begin{defn}
+{\upshape\cite[Page 495 (Cartesian product)]{Blackburn}}
+\begin{holmath}
+  \HOLConst{Cart_prod}\;\HOLFreeVar{J}\;\HOLFreeVar{\ensuremath{A\sp{\sf s}}}\;\HOLTokenDefEquality{}\;\HOLTokenLeftbrace{}\HOLBoundVar{f}\;\HOLTokenBar{}\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{j}.\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLBoundVar{f}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{A\sp{\sf s}}}\;\HOLBoundVar{j}\HOLTokenRightbrace{}
+\end{holmath}
+\end{defn}
+
+If $U$ is an ultrafilter on $J$, for two functions $f,g$ in the Cartesian product \HOLinline{\HOLConst{Cart_prod}\;\HOLFreeVar{J}\;\HOLFreeVar{\ensuremath{A\sp{\sf s}}}}, we say $f$ and $g$ are $U$-equivalent (notation: \HOLinline{\HOLFreeVar{f}\sim\sb{\HOLFreeVar{U}}\sp{\HOLFreeVar{\ensuremath{A\sp{\sf s}}}}\HOLFreeVar{g}}) if the set \HOLinline{\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{f}\;\HOLBoundVar{j}\;\HOLSymConst{=}\;\HOLFreeVar{g}\;\HOLBoundVar{j}\HOLTokenRightbrace{}} (where the values of $f$ and $g$ agree) is in $U$.
+The ultraproduct of  $(A_j)_{j\in J}$ modulo $U$ is the quotient of \HOLinline{\HOLConst{Cart_prod}\;\HOLFreeVar{J}\;\HOLFreeVar{\ensuremath{A\sp{\sf s}}}} by \HOLinline{\sim\sb{\HOLFreeVar{U}}\sp{\HOLFreeVar{\ensuremath{A\sp{\sf s}}}}}.
+\begin{defn}
+{\upshape\cite[Definition 2.69 (Ultraproduct of Sets)]{Blackburn}}
+\begin{holmath}
+  \HOLConst{ultraproduct}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;\HOLFreeVar{\ensuremath{A\sp{\sf s}}}\;\HOLTokenDefEquality{}\;\HOLConst{Cart_prod}\;\HOLFreeVar{J}\;\HOLFreeVar{\ensuremath{A\sp{\sf s}}}\HOLSymConst{/}\sim\sb{\HOLFreeVar{U}}\sp{\HOLFreeVar{\ensuremath{A\sp{\sf s}}}}
+\end{holmath}
+\end{defn}
+We write \HOLinline{\HOLFreeVar{\ensuremath{f\sb{U}}}} to denote the equivalence class that $f$ belongs to.
+In the case where \HOLinline{\HOLFreeVar{\ensuremath{A\sp{\sf s}}}\;\HOLFreeVar{j}\;\HOLSymConst{=}\;\HOLFreeVar{A}} for all $j\in J$, the ultraproduct is called the ultrapower of \HOLinline{\HOLFreeVar{A}} modulo $U$.
+As before, in the definition above, the family $(A_j)_{j\in J}$ is encoded as a function, and hence for $j\in J$, \HOLinline{\HOLFreeVar{\ensuremath{A\sp{\sf s}}}\;\HOLFreeVar{j}} is the set $A_j$ indexed by $j$.
+\subsubsection{Ultraproduct for modal models}
+Given a family \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}} of modal models indexed by $J$ and an ultrafilter $U$ on $J$, the ultraproduct model of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}} modulo $U$ (notation : \HOLinline{\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}}) is described as follows:
+
+\begin{itemize}
+  \item The world set is the ultraproduct of world sets of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}} modulo $U$.
+
+  \item Two equivalence classes $f_U,g_U$ of functions are related in \HOLinline{\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}} iff there exist $f_0\in f_U,g_0\in g_U$, such that \HOLinline{\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLTokenBar{}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j})\HOLSymConst{\ensuremath{\sp{R}}}\;(\HOLFreeVar{f}\sb{\mathrm{0}}\;\HOLBoundVar{j})\;(\HOLFreeVar{g}\sb{\mathrm{0}}\;\HOLBoundVar{j})\HOLTokenRightbrace{}} is in $U$.
+  \item A propositional letter $p$ is satisfied at $f_U$ in \HOLinline{\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}} iff there exists $f_0\in f_U$ such that \HOLinline{\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{f}\sb{\mathrm{0}}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j})\HOLSymConst{\ensuremath{\sp{V}}}\;\HOLFreeVar{p}\HOLTokenRightbrace{}} is in $U$.
+\end{itemize}
+\begin{defn}
+{\upshape\cite[Definition 2.70 (Ultraproduct of Modal Models)]{Blackburn}}
+\begin{holmath}
+  \HOLConst{ultraproduct_model}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLTokenDefEquality{}\\
+\;\;\HOLTokenLeftrec{}\HOLFieldName{frame}\;:=\\
+\;\;\;\;\HOLTokenLeftrec{}\HOLFieldName{world}\;:=\;\HOLConst{ultraproduct}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;(\HOLConst{worlds}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}});\\
+\;\;\;\;\;\;\HOLFieldName{rel}\;:=\\
+\;\;\;\;\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLBoundVar{\ensuremath{g\sb{U}}}.\\
+\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{f}\sb{\mathrm{0}}\;\HOLBoundVar{g}\sb{\mathrm{0}}.\\
+\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLBoundVar{f}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{g}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{g\sb{U}}}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j})\HOLSymConst{\ensuremath{\sp{R}}}\;(\HOLBoundVar{f}\sb{\mathrm{0}}\;\HOLBoundVar{j})\;(\HOLBoundVar{g}\sb{\mathrm{0}}\;\HOLBoundVar{j})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U})\HOLTokenRightrec{};\\
+\;\;\HOLFieldName{valt}\;:=\\
+\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{p}\;\HOLBoundVar{\ensuremath{f\sb{U}}}.\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{f}\sb{\mathrm{0}}.\;\HOLBoundVar{f}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{f}\sb{\mathrm{0}}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j})\HOLSymConst{\ensuremath{\sp{V}}}\;\HOLBoundVar{p}\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U})\HOLTokenRightrec{}
+\end{holmath}
+\end{defn}
+
+Here \HOLinline{\HOLConst{worlds}} is the function that takes a family of models to the family of their world sets:
+\begin{holmath}
+  \HOLConst{worlds}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLTokenDefEquality{}\;(\HOLTokenLambda{}\HOLBoundVar{j}.\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j})\HOLSymConst{\ensuremath{\sp{W}}})
+\end{holmath}
+
+As \HOLinline{\sim\sb{\HOLFreeVar{U}}\sp{\HOLFreeVar{A}}} is an equivalence relation, if one element in an equivalence class satisfies the required condition, then all the elements in the equivalence class will satisfy the condition.
+Therefore, if we replace all the existential quantifiers with universal quantifiers in the above definition, the construction is still valid, and will give the same model as the current definition.
+
+The critical result we will need about ultraproducts of modal models is a modal version of the fundamental theorem of ultraproducts, which is also called \L o\'s's theorem.
+
+\begin{thm}[\texttt{Los_modal_thm}]
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{ultrafilter}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{f\sb{U}}}\;\HOLSymConst{\HOLTokenIn{}}\;(\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}})\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;(\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}},\HOLFreeVar{\ensuremath{f\sb{U}}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{f}\sb{\mathrm{0}}.\;\HOLBoundVar{f}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{f\sb{U}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j},\HOLBoundVar{f}\sb{\mathrm{0}}\;\HOLBoundVar{j}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U})
+\end{holmath}
+\end{thm}
+Although it is possible to derive the theorem above from \L o\'s's theorem of first-order models using standard translation. We proved it directly by induction on modal formulas in HOL.
+
+\subsubsection{Ultraproduct for first-order models}
+Given a family \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}} of first-order models indexed by $J$ and an ultrafilter $U$ on $J$, the ultraproduct model of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}} modulo $U$ (notation : \HOLinline{{}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}}) is given by:
+
+\begin{itemize}
+\item The domain is the ultraproduct of the domains of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}} over $U$ on $J$.
+\item A function with its symbol denoted by the natural number $n$ will send a list \HOLinline{\HOLFreeVar{zs}} of equivalence classes to the equivalence class of a function that sending $j\in J$ to \HOLinline{(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLFreeVar{j}).\HOLFieldName{Fun}\;\HOLFreeVar{n}\;\HOLFreeVar{l}}, where the $k$-th member of the list $l$ is a representative of the $k$-th member (which is an equivalence class) of \HOLinline{\HOLFreeVar{zs}}.
+
+\item A predicate with its symbol denoted by $p$ will hold for a list $zs$ of equivalence classes if and only if once we have a list $zr$ such that the $k$-th member is a representative of the $k$-th member of $zs$, the set of elements \HOLinline{\HOLFreeVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}} such that \HOLinline{(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLFreeVar{j}).\HOLFieldName{Pred}\;\HOLFreeVar{p}\;\HOLFreeVar{zr}} is in $U$.
+\end{itemize}
+\begin{defn}
+{\upshape\cite[Definition A.18 (Ultraproduct of First-Order Models)]{Blackburn}}
+\begin{holmath}
+  {}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLTokenDefEquality{}\\
+\;\;\HOLTokenLeftrec{}\HOLFieldName{Dom}\;:=\;\HOLConst{ultraproduct}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;(\HOLConst{Doms}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}});\\
+\;\;\HOLFieldName{Fun}\;:=\\
+\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{n}\;\HOLBoundVar{zs}.\\
+\;\;\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLBoundVar{y}\;\HOLTokenBar{}\\
+\;\;\;\;\;\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{j}.\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLBoundVar{y}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Dom})\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{y}\;\HOLBoundVar{j}\;\HOLSymConst{=}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Fun}\;\HOLBoundVar{n}\;(\HOLConst{MAP}\;(\HOLTokenLambda{}\HOLBoundVar{\ensuremath{f\sb{U}}}.\;\HOLConst{CHOICE}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLBoundVar{j})\;\HOLBoundVar{zs})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U}\HOLTokenRightbrace{});\\
+\;\;\HOLFieldName{Pred}\;:=\;(\HOLTokenLambda{}\HOLBoundVar{p}\;\HOLBoundVar{zs}.\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Pred}\;\HOLBoundVar{p}\;(\HOLConst{MAP}\;(\HOLTokenLambda{}\HOLBoundVar{\ensuremath{f\sb{U}}}.\;\HOLConst{CHOICE}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLBoundVar{j})\;\HOLBoundVar{zs})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U})\HOLTokenRightrec{}
+\end{holmath}
+\end{defn}
+Here we fix the representative of each equivalence class \HOLinline{\HOLFreeVar{\ensuremath{f\sb{U}}}} to be \HOLinline{\HOLConst{CHOICE}\;\HOLFreeVar{\ensuremath{f\sb{U}}}}.
+The function \HOLinline{\HOLConst{Doms}} takes a family of first-order models to the family of their domains.
+
+The semantic behavior of ultraproduct models are characterized by \emph{\L o\'s's theorem}, whose proof can be founded in \cite{ChangKeisler1990}.
+
+\begin{thm}
+{\upshape\cite[Theorem A.19 (\L o\'s's theorem)]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{ultrafilter}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{valuation}\;({}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}})\;\HOLFreeVar{\ensuremath{\sigma}}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{j}.\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{wffm}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}))\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLConst{termval}\;({}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}})\;\HOLFreeVar{\ensuremath{\sigma}}\;\HOLFreeVar{t}\;\HOLSymConst{=}\\
+\;\;\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLBoundVar{f}\;\HOLTokenBar{}\;\HOLBoundVar{f}\sim\sb{\HOLFreeVar{U}}\sp{\HOLConst{Doms}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}}(\HOLTokenLambda{}\HOLBoundVar{j}.\;\HOLConst{termval}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j})\;(\HOLTokenLambda{}\HOLBoundVar{v}.\;\HOLConst{CHOICE}\;(\HOLFreeVar{\ensuremath{\sigma}}\;\HOLBoundVar{v})\;\HOLBoundVar{j})\;\HOLFreeVar{t})\HOLTokenRightbrace{} \\[3mm]
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{ultrafilter}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{valuation}\;({}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}})\;\HOLFreeVar{\ensuremath{\sigma}}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{j}.\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{wffm}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}))\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;({}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}},\HOLFreeVar{\ensuremath{\sigma}}\ensuremath{\vDash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenEquiv{}}\\
+\;\;\;\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j},(\HOLTokenLambda{}\HOLBoundVar{v}.\;\HOLConst{CHOICE}\;(\HOLFreeVar{\ensuremath{\sigma}}\;\HOLBoundVar{v})\;\HOLBoundVar{j})\ensuremath{\vDash}\;\HOLFreeVar{\ensuremath{\phi}}\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U})
+\end{holmath}
+\end{thm}
+
+\section{Ultrafilter Extensions}[2p]
+The first application of the theory of ultrafilters above is to construct the ultrafilter extension of a model, which has the nice property of being M-saturated.
+To define M-saturation, we give the following three definitions (the first two are called \emph{finitely satisfiable}, \emph{satisfiable}) consecutively:
+\begin{defn}
+{\upshape\cite[Definition 2.53]{Blackburn}} 
+\begin{holmath}
+  \HOLConst{fin_satisfiable_in}\;\HOLFreeVar{\ensuremath{\Sigma}}\;\HOLFreeVar{X}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLTokenDefEquality{}\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{S}.\;\HOLBoundVar{S}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\Sigma}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{S}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{satisfiable_in}\;\HOLBoundVar{S}\;\HOLFreeVar{X}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\\[3mm]
+  \HOLConst{satisfiable_in}\;\HOLFreeVar{\ensuremath{\Sigma}}\;\HOLFreeVar{X}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLTokenDefEquality{}\;\HOLFreeVar{X}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{w}.\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{X}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{\ensuremath{\phi}}.\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\Sigma}}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLBoundVar{\ensuremath{\phi}}\\[3mm]
+  \HOLConst{M_sat}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLTokenDefEquality{}\\
+\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{w}\;\HOLBoundVar{\ensuremath{\Sigma}}.\\
+\;\;\;\;\;\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{fin_satisfiable_in}\;\HOLBoundVar{\ensuremath{\Sigma}}\;\HOLTokenLeftbrace{}\HOLBoundVar{v}\;\HOLTokenBar{}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{R}}}\;\HOLBoundVar{w}\;\HOLBoundVar{v}\HOLTokenRightbrace{}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\;\HOLConst{satisfiable_in}\;\HOLBoundVar{\ensuremath{\Sigma}}\;\HOLTokenLeftbrace{}\HOLBoundVar{v}\;\HOLTokenBar{}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{R}}}\;\HOLBoundVar{w}\;\HOLBoundVar{v}\HOLTokenRightbrace{}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}
+\end{holmath}
+\end{defn}
+For M-saturated models, bisimulation and modal equivalence coincides:
+\begin{prop}\label{2.54}
+{\upshape\cite[Proposition 2.54]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{M_sat}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{M_sat}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{w}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\sb{\mathrm{1}}\ensuremath{\leftrightsquigarrow}\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLFreeVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\HOLSymConst{,}\HOLFreeVar{w}\sb{\mathrm{1}}\HOLSymConst{\ensuremath{\leftrightarroweq}}\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{,}\HOLFreeVar{w}\sb{\mathrm{2}}
+\end{holmath}
+\end{prop}
+Given a model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}}, its ultrafilter extension \HOLinline{{}\sp{ue}{\HOLFreeVar{\ensuremath{\mathfrak{M}}}}} is defined as:
+\begin{defn}
+{\upshape\cite[Definition 2.57 (Ultrafilter Extension)]{Blackburn}}
+\begin{holmath}
+  {}\sp{ue}{\HOLFreeVar{\ensuremath{\mathfrak{M}}}}\;\HOLTokenDefEquality{}\\
+\;\;\HOLTokenLeftrec{}\HOLFieldName{frame}\;:=\\
+\;\;\;\;\HOLTokenLeftrec{}\HOLFieldName{world}\;:=\;\HOLTokenLeftbrace{}\HOLBoundVar{u}\;\HOLTokenBar{}\;\HOLConst{ultrafilter}\;\HOLBoundVar{u}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\HOLTokenRightbrace{};\\
+\;\;\;\;\;\;\HOLFieldName{rel}\;:=\\
+\;\;\;\;\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{u}\;\HOLBoundVar{v}.\\
+\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLConst{ultrafilter}\;\HOLBoundVar{u}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{ultrafilter}\;\HOLBoundVar{v}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X}.\;\HOLBoundVar{X}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenImp{}}\;{\HOLFreeVar{\ensuremath{\mathfrak{M}}}}\sb{\Diamond}(\HOLBoundVar{X})\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{u})\HOLTokenRightrec{};\\
+\;\;\HOLFieldName{valt}\;:=\;(\HOLTokenLambda{}\HOLBoundVar{p}\;\HOLBoundVar{v}.\;\HOLConst{ultrafilter}\;\HOLBoundVar{v}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLTokenLeftbrace{}\HOLBoundVar{w}\;\HOLTokenBar{}\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{V}}}\;\HOLBoundVar{p}\;\HOLBoundVar{w}\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{v})\HOLTokenRightrec{}
+\end{holmath}
+\end{defn}
+
+Using the ultrafilter theorem and some basic properties about ultrafilters, we derive:
+ \begin{prop}\label{2.59}
+{\upshape\cite[Proposition 2.59 (i)]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{ultrafilter}\;\HOLFreeVar{u}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;(\HOLTokenLeftbrace{}\HOLBoundVar{w}\;\HOLTokenBar{}\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{u}\;\HOLSymConst{\HOLTokenEquiv{}}\;{}\sp{ue}{\HOLFreeVar{\ensuremath{\mathfrak{M}}}},\HOLFreeVar{u}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}})
+\end{holmath}
+\end{prop}
+In particular, every world \HOLinline{\HOLFreeVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}} is embedded as the principal filter \HOLinline{\pi\sb{\HOLFreeVar{w}}\sp{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}}} on \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}} generated by \HOLinline{\HOLFreeVar{w}} in the ultrafilter extension or \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}}.
+Also, the above leads to the proof of the fact that the ultrafilter extension of every model is M-saturated.
+The M-saturatedness of ultrafilter extensions together with \ref{2.54} immediately gives the central result about ultrafilter extension: bisimularity of worlds in a model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} can be characterized as bisimularity in \HOLinline{{}\sp{ue}{\HOLFreeVar{\ensuremath{\mathfrak{M}}}}}.
+\begin{thm}
+{\upshape\cite[Proposition 2.62]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLFreeVar{w}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\sb{\mathrm{1}}\ensuremath{\leftrightsquigarrow}\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLFreeVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenEquiv{}}\;{}\sp{ue}{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}}\HOLSymConst{,}\pi\sb{\HOLFreeVar{w}\sb{\mathrm{1}}}\sp{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\HOLSymConst{\ensuremath{\sp{W}}}}\HOLSymConst{\ensuremath{\leftrightarroweq}}{}\sp{ue}{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}}\HOLSymConst{,}\pi\sb{\HOLFreeVar{w}\sb{\mathrm{2}}}\sp{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}})
+\end{holmath}
+\end{thm}
+
+\section{Countably Saturatedness of Ultrapower Models}[2p]
+
+Given a first-order model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} with no information about interpretation of function symbols, we can expand the model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} by adding the interpretation of some function symbols.
+For our propose, we are only interested in adding the interpretation of finitely many nullary function symbols, which are also called \emph{constants}.
+We write \HOLinline{\HOLConst{is_expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}} to mean that \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}} is the result of adding each element in $A$ to \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} as a new constant.
+Further, the function $f$ is a bijection between $\{0,\cdots,n-1\}$ and $A$, which is assumed to be finite, so that each nullary function symbol $c$ will be interpreted as \HOLinline{\HOLFreeVar{f}\;\HOLFreeVar{c}} in \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}}.
+\begin{defn}
+{\upshape\cite[Definition A.9 (Expansion)]{Blackburn}}
+\begin{holmath}
+  \HOLConst{is_expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}\;\HOLTokenDefEquality{}\\
+\;\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}.\HOLFieldName{Dom}\;\HOLSymConst{=}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{BIJ}\;\HOLFreeVar{f}\;(\HOLConst{count}\;(\HOLConst{CARD}\;\HOLFreeVar{A}))\;\HOLFreeVar{A}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}.\HOLFieldName{Fun}\;\HOLSymConst{=}\\
+\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{c}\;\HOLBoundVar{l}.\;\HOLKeyword{if}\;\HOLBoundVar{c}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLConst{CARD}\;\HOLFreeVar{A}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{l}\;\HOLSymConst{=}\;[]\;\HOLKeyword{then}\;\HOLFreeVar{f}\;\HOLBoundVar{c}\;\HOLKeyword{else}\;\HOLConst{CHOICE}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom})\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}.\HOLFieldName{Pred}\;\HOLSymConst{=}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Pred}
+\end{holmath}
+\end{defn}
+If \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}}, \HOLinline{\HOLFreeVar{A}} and \HOLinline{\HOLFreeVar{f}} are all fixed, then the model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}} such that \HOLinline{\HOLConst{is_expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}} is unique.
+The only difference between a model and an expansion of it is the interpretation of function symbols.
+
+A set \HOLinline{\HOLFreeVar{\ensuremath{\Sigma}}} of first-order formulas is called \emph{consistent} with a model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} if for every finite subset \HOLinline{\HOLFreeVar{\ensuremath{\Sigma}}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\Sigma}}}, there exists a valuation of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} such that all elements of \HOLinline{\HOLFreeVar{\ensuremath{\Sigma}}\sb{\mathrm{0}}} are satisfied, in this case, we write \HOLinline{\HOLConst{consistent}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{\ensuremath{\Sigma}}}.
+A set \HOLinline{\HOLFreeVar{\ensuremath{\Gamma}}} of first-order formula is an $x$-\emph{type} if for each formula in \HOLinline{\HOLFreeVar{\ensuremath{\Gamma}}}, the only free variable that may contain is $x$.
+In this case, we write `\HOLinline{\HOLConst{ftype}\;\HOLFreeVar{x}\;\HOLFreeVar{\ensuremath{\Gamma}}}' in HOL.
+If \HOLinline{\HOLFreeVar{\ensuremath{\Gamma}}} is an $x$-type, when evaluating formulas in \HOLinline{\HOLFreeVar{\ensuremath{\Gamma}}}, the valuations will only control where the only free variable $x$ goes to.
+We say \HOLinline{\HOLFreeVar{\ensuremath{\Gamma}}} is \emph{realized} in \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} if there is an element $w$ in the domain of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} such that \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}},(\HOLTokenLambda{}\HOLBoundVar{v}.\;\HOLFreeVar{w})\ensuremath{\vDash}\;\HOLFreeVar{\ensuremath{\phi}}} for all \HOLinline{\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\Gamma}}}.
+In this case, we write `\HOLinline{\HOLConst{frealizes}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{x}\;\HOLFreeVar{\ensuremath{\Gamma}}}' in HOL.
+Let \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} be a model and \HOLinline{\HOLFreeVar{n}} be a natural number. 
+%
+For every \HOLinline{\HOLFreeVar{A}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}}, with $|A|<n$ and for every $f:{\mathbb N}\to $ \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}}, there is a unique \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}} such that  \HOLinline{\HOLConst{is_expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}}.
+%
+If every such \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}} realizes every $x$-type \HOLinline{\HOLFreeVar{\ensuremath{\Gamma}}}, then we say \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} is $n$-saturated.
+In HOL:
+\begin{defn}
+{\upshape\cite[Definition 2.63 ($n$-Saturated)]{Blackburn}}
+\begin{holmath}
+  \HOLConst{n_saturated}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{n}\;\HOLTokenDefEquality{}\\
+\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{A}\;\HOLBoundVar{\ensuremath{\mathfrak{M}'}}\;\HOLBoundVar{\ensuremath{\Gamma}}\;\HOLBoundVar{x}\;\HOLBoundVar{f}.\\
+\;\;\;\;\;\;\HOLConst{IMAGE}\;\HOLBoundVar{f}\;\ensuremath{{\cal{U}}}({:}\HOLTyOp{num})\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{A}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{CARD}\;\HOLBoundVar{A}\;\HOLSymConst{\HOLTokenLeq{}}\;\HOLFreeVar{n}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{A}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\HOLConst{is_expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLBoundVar{A}\;\HOLBoundVar{\ensuremath{\mathfrak{M}'}}\;\HOLBoundVar{f}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{\ensuremath{\phi}}.\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\Gamma}}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{form_functions}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLTokenLeftbrace{}(\HOLBoundVar{c}\HOLSymConst{,}\HOLNumLit{0})\;\HOLTokenBar{}\;\HOLBoundVar{c}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLConst{CARD}\;\HOLBoundVar{A}\HOLTokenRightbrace{})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{ftype}\;\HOLBoundVar{x}\;\HOLBoundVar{\ensuremath{\Gamma}}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\HOLConst{consistent}\;\HOLBoundVar{\ensuremath{\mathfrak{M}'}}\;\HOLBoundVar{\ensuremath{\Gamma}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\;\HOLConst{frealizes}\;\HOLBoundVar{\ensuremath{\mathfrak{M}'}}\;\HOLBoundVar{x}\;\HOLBoundVar{\ensuremath{\Gamma}}\\[3mm]
+\end{holmath}
+\end{defn}
+We say \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} is countably saturated, and write \HOLinline{\HOLConst{countably_saturated}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}} if \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} is $n$-saturated for every natural number $n$.
+The ultimate goal is to prove a lemma to be used in the proof of Van Benthem characterization theorem: For a family of non-empty models, their ultraproduct on a countably incomplete ultrafilter is \emph{countably saturated}.
+\begin{lm}\label{2.73}
+{\upshape\cite[Lemma 2.73]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{countably_incomplete}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{j}.\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenImp{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j})\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLSymConst{\HOLTokenEmpty{}})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLConst{countably_saturated}\;(\HOLConst{mm2folm}\;(\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}))
+\end{holmath}
+\end{lm}
+Here a countably incomplete ultrafilter is an ultrafilter that contains a countably infinite family that intersects to the empty set.
+We proved in HOL that such ultrafilters do exists using \ref{A.14}.
+The above theorem is not simply a direct consequence from the \L o\'s's theorem:
+\L o\'s's theorem is about ultraproducts of first-order models, and it says nothing about expansion.
+But to prove \ref{2.73}, we are required to prove a statement for a model obtained by expanding a first-order model which is again obtained by viewing an ultraproduct of modal models as a first-order model.
+
+To deal with this issue, the key observation is that constants are nothing more than forcing some symbols to be sent to some points in a model under every valuation, hence rather than use nullary function symbols, we fixed a set of variable letters, each corresponds to a function symbol, and only consider the valuations that sends these variable letters to fixed certain points.
+With this idea, we can remove all the constants in a formula, and hence change our scope from an expanded model back to the unexpanded model.
+To get rid of the constants $\{0,\cdots , n-1\}$, we replace every \HOLinline{\HOLConst{VAR}\;\HOLFreeVar{m}} with \HOLinline{\HOLConst{VAR}\;(\HOLFreeVar{m}\;\HOLSymConst{\ensuremath{+}}\;\HOLFreeVar{n})}, and replace every constant \HOLinline{\HOLConst{\ensuremath{{}^{\sf f}{\sf Fn}}}\;\HOLFreeVar{c}\;[]} by \HOLinline{\HOLConst{VAR}\;\HOLFreeVar{c}}.
+This operation is done by the function \HOLinline{\HOLConst{shift_form}} which takes a natural number (the number of constants we want to remove), and a first-order formula (where the only function symbols may appear are the constants $0,\cdots, n-1$).
+Since $0,\cdots,n-1$ in a shifted formula are now designed to be sent to fixed places $f \ 0,\cdots,f \ (n-1)$, it does not make sense to assign these variable symbols anywhere else.
+Therefore, to talk about evaluation of shifted formula, the first thing is to make sure that the valuations we are considering send the variables which actually denotes constants to the right place.
+Hence we shift the valuations accordingly.
+The definition of shifting on valuation together with the semantical behaviour are displayed below:
+\begin{defn}[Shifting on valuations, \texttt{expansion_shift_feval}]\label{k1}
+\begin{holmath}
+  \HOLConst{shift_valuation}\;\HOLFreeVar{n}\;\HOLFreeVar{\ensuremath{\sigma}}\;\HOLFreeVar{f}\;\HOLTokenDefEquality{}\;(\HOLTokenLambda{}\HOLBoundVar{v}.\;\HOLKeyword{if}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLFreeVar{n}\;\HOLKeyword{then}\;\HOLFreeVar{f}\;\HOLBoundVar{v}\;\HOLKeyword{else}\;\HOLFreeVar{\ensuremath{\sigma}}\;(\HOLBoundVar{v}\;\HOLSymConst{\ensuremath{-}}\;\HOLFreeVar{n}))\\[2mm]
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{is_expansion}\;(\HOLConst{mm2folm}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}})\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{valuation}\;(\HOLConst{mm2folm}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}})\;\HOLFreeVar{\ensuremath{\sigma}}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\HOLConst{form_functions}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLTokenLeftbrace{}(\HOLBoundVar{c}\sb{\mathrm{1}}\HOLSymConst{,}\HOLNumLit{0})\;\HOLTokenBar{}\;\HOLBoundVar{c}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLConst{CARD}\;\HOLFreeVar{A}\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;(\HOLFreeVar{\ensuremath{\mathfrak{M}'}},\HOLFreeVar{\ensuremath{\sigma}}\ensuremath{\vDash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenEquiv{}}\\
+\;\;\;\;\;\;\;\;\;\;\HOLConst{mm2folm}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLConst{shift_valuation}\;(\HOLConst{CARD}\;\HOLFreeVar{A})\;\HOLFreeVar{\ensuremath{\sigma}}\;\HOLFreeVar{f}\ensuremath{\vDash}\;\HOLConst{shift_form}\;(\HOLConst{CARD}\;\HOLFreeVar{A})\;\HOLFreeVar{\ensuremath{\phi}})
+\end{holmath}
+\end{defn}
+
+The shifting construction gets us out of the expansion, leaving us a model obtained by converting a ultraproduct modal model to a first-order model.
+To apply \L o\'s's theorem on such a model, we prove by induction that:
+\begin{prop}[\texttt{ultraproduct_comm_feval}]\label{k2}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{ultrafilter}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{form_functions}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{=}\;\HOLSymConst{\HOLTokenEmpty{}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{valuation}\;(\HOLConst{mm2folm}\;(\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}))\;\HOLFreeVar{\ensuremath{\sigma}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;(\HOLConst{mm2folm}\;(\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}),\HOLFreeVar{\ensuremath{\sigma}}\ensuremath{\vDash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenEquiv{}}\;{}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;(\HOLTokenLambda{}\HOLBoundVar{j}.\;\HOLConst{mm2folm}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j})),\HOLFreeVar{\ensuremath{\sigma}}\ensuremath{\vDash}\;\HOLFreeVar{\ensuremath{\phi}})
+\end{holmath}
+\end{prop}
+By \ref{k1} and \ref{k2}, proving \ref{2.73} amounts to prove: 
+\begin{lm}[Saturation of ultraproduct model, \texttt{ultraproduct_sat}]
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{countably_incomplete}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{valuation}\;({}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}})\;\HOLFreeVar{f}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{j}.\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{wffm}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}))\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{\ensuremath{\phi}}.\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\Delta}}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\ensuremath{{\mathcal L}^1_\tau}}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FV}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLConst{\ensuremath{\setminus}}\;\HOLFreeVar{C}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLTokenLeftbrace{}\HOLFreeVar{x}\HOLTokenRightbrace{})\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{\ensuremath{\Delta}}\sb{\mathrm{0}}.\\
+\;\;\;\;\;\;\;\;\;\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\Delta}}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\Delta}}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\Delta}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{\ensuremath{\sigma}}.\\
+\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLConst{valuation}\;({}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}})\;\HOLBoundVar{\ensuremath{\sigma}}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{c}.\;\HOLBoundVar{c}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{C}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLBoundVar{\ensuremath{\sigma}}\;\HOLBoundVar{c}\;\HOLSymConst{=}\;\HOLFreeVar{f}\;\HOLBoundVar{c})\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{\ensuremath{\phi}}.\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\Delta}}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenImp{}}\;{}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}},\HOLBoundVar{\ensuremath{\sigma}}\ensuremath{\vDash}\;\HOLBoundVar{\ensuremath{\phi}})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{\ensuremath{\sigma}}.\\
+\;\;\;\;\;\;\;\;\;\;\;\HOLConst{valuation}\;({}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}})\;\HOLBoundVar{\ensuremath{\sigma}}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{c}.\;\HOLBoundVar{c}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{C}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLBoundVar{\ensuremath{\sigma}}\;\HOLBoundVar{c}\;\HOLSymConst{=}\;\HOLFreeVar{f}\;\HOLBoundVar{c})\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{\ensuremath{\phi}}.\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\Delta}}\;\HOLSymConst{\HOLTokenImp{}}\;{}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}},\HOLBoundVar{\ensuremath{\sigma}}\ensuremath{\vDash}\;\HOLBoundVar{\ensuremath{\phi}}
+\end{holmath}
+\end{lm}
+The proof of the above lemma requires rather heavy construction and can be found in \cite{ChangKeisler1990}.
+
+\section{Van Benthem's Characterization Theorem}[2p]
+Note that the standard translation of any modal formula only can only contain unary predicate symbols which corresponds to propositional letters, one binary predicate symbol which corresponds to the relation, and no function symbol.
+A first-order formula which only use these symbols is called an ${\mathcal L}_{\tau}^1$-formula.
+An ${\mathcal L}_{\tau}^1$-formula which contains only one free variable is called \emph{invariant under bisimulation} if for all models \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} and \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{N}}}} with \HOLinline{\HOLFreeVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}} and \HOLinline{\HOLFreeVar{v}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{N}}}\HOLSymConst{\ensuremath{\sp{W}}}}, if there exists a bisimulation relation between \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} and \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{N}}}} relating $w$ and $v$, then $\phi$ holds at \HOLinline{\HOLFreeVar{w}} if and only if it holds at \HOLinline{\HOLFreeVar{v}} when both \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} and \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{N}}}} are viewed as first-order models.
+\begin{defn}
+{\upshape\cite[Definition 2.67 (Invariant for Bisimulations)]{Blackburn}}
+\begin{holmath}
+  \HOLConst{invar4bisim}\;(\HOLFreeVar{x} :\HOLTyOp{num})\;({:}\ensuremath{\alpha})\;({:}\ensuremath{\beta})\;(\HOLFreeVar{\ensuremath{\phi}} :\HOLTyOp{folform})\;\HOLTokenDefEquality{}\\
+\;\;\HOLConst{FV}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLTokenLeftbrace{}\HOLFreeVar{x}\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\ensuremath{{\mathcal L}^1_\tau}}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :(\HOLTyOp{num},\;\ensuremath{\alpha})\;\HOLTyOp{model})\;(\HOLBoundVar{\ensuremath{\mathfrak{N}}} :(\HOLTyOp{num},\;\ensuremath{\beta})\;\HOLTyOp{model})\;(\HOLBoundVar{v} :\ensuremath{\beta})\;(\HOLBoundVar{w} :\ensuremath{\alpha}).\\
+\;\;\;\;\;\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{,}\HOLBoundVar{w}\HOLSymConst{\ensuremath{\leftrightarroweq}}\HOLBoundVar{\ensuremath{\mathfrak{N}}}\HOLSymConst{,}\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\;(\HOLConst{mm2folm}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},(\HOLTokenLambda{}(\HOLBoundVar{x} :\HOLTyOp{num}).\;\HOLBoundVar{w})\ensuremath{\vDash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenEquiv{}}\\
+\;\;\;\;\;\;\;\;\;\;\;\HOLConst{mm2folm}\;\HOLBoundVar{\ensuremath{\mathfrak{N}}},(\HOLTokenLambda{}(\HOLBoundVar{x} :\HOLTyOp{num}).\;\HOLBoundVar{v})\ensuremath{\vDash}\;\HOLFreeVar{\ensuremath{\phi}})
+\end{holmath}
+\end{defn}
+
+By the same reason as ??, the type parameters here is necessasy.
+Although it is possible to prove theorems for different types $\alpha$ and $\beta$ in the above definition, we will only consider the case that $\alpha$ and $\beta$ are the same when proving theorems.
+
+The Van Benthem characterization theorem says an $\mathcal L_{\tau}^1$ formula with at most one free variable $x$ is invariant under bisimulation precisely when it is equivalent to the standard translation of some modal formula at $x$.
+It is immediate from \ref{2.47} that every such formula which is equivalent to a standard translation is invariant for bisimulation.
+We cannot prove it as an `if and only if' statement, since according to the proofs in \cite{Blackburn}, we can only prove the two directions separately as:
+\begin{prop}
+{\upshape\cite[Theorem 2.68, as two separate directions]{Blackburn}}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{FV}\;(\HOLFreeVar{\ensuremath{\delta}} :\HOLTyOp{folform})\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLTokenLeftbrace{}(\HOLFreeVar{x} :\HOLTyOp{num})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\ensuremath{{\mathcal L}^1_\tau}}\;\HOLFreeVar{\ensuremath{\delta}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\delta}}\;\;{}\sp{\sf f}{\equiv}\sb{({:}\ensuremath{\alpha})}\;\;{\sf ST}\sb{\HOLFreeVar{x}}\;(\HOLFreeVar{\ensuremath{\phi}} :\HOLTyOp{num}\;\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLConst{invar4bisim}\;\HOLFreeVar{x}\;({:}\ensuremath{\alpha})\;({:}\ensuremath{\alpha})\;\HOLFreeVar{\ensuremath{\delta}} \\[3mm]
+ \ensuremath{\HOLTokenTurnstile}\HOLConst{INFINITE}\;\ensuremath{{\cal{U}}}({:}\ensuremath{\alpha})\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\HOLConst{invar4bisim}\;(\HOLFreeVar{x} :\HOLTyOp{num})\;({:}(\HOLTyOp{num}\;\ensuremath{\rightarrow}\;\ensuremath{\alpha})\;\ensuremath{\rightarrow}\;\HOLTyOp{bool})\;({:}(\HOLTyOp{num}\;\ensuremath{\rightarrow}\;\ensuremath{\alpha})\;\ensuremath{\rightarrow}\;\HOLTyOp{bool})\\
+\;\;\;\;\;\;\;(\HOLFreeVar{\ensuremath{\delta}} :\HOLTyOp{folform})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\phi}} :\HOLTyOp{num}\;\HOLTyOp{form}).\;\HOLFreeVar{\ensuremath{\delta}}\;\;{}\sp{\sf f}{\equiv}\sb{({:}\ensuremath{\alpha})}\;\;{\sf ST}\sb{\HOLFreeVar{x}}\;\HOLBoundVar{\ensuremath{\phi}}
+\end{holmath}
+\end{prop}
+
+which cannot be put together into a double implication.
+To see the reason: given an $\mathcal L_{\tau}^1$-formula $\phi$ with no more then one free variable, by the second theorem above, if \HOLinline{\HOLFreeVar{\ensuremath{\phi}}} is invariant under bisimulation for models with \HOLinline{(\HOLTyOp{num}\;\ensuremath{\rightarrow}\;\ensuremath{\alpha})\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}}-worlds, then $\phi$ is equivalent to a standard translation on model with $\alpha$-worlds.
+However, if we want to prove the converse of this statement, we need to start with the assumption that $\phi$ is equivalent to a standard translation on models with $\alpha$-worlds, and prove that $\phi$ is invariant for bisimulation for models with \HOLinline{(\HOLTyOp{num}\;\ensuremath{\rightarrow}\;\ensuremath{\alpha})\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}}-world.
+But by the first theorem above, we can only conclude $\phi$ is invariant for bisimulation for models of type $\alpha$.
+If the type universe of \HOLinline{(\HOLTyOp{num}\;\ensuremath{\rightarrow}\;\ensuremath{\alpha})\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}} is small enough to be embedded into $\alpha$, then we will also done.
+However, the cardinality of the universe of \HOLinline{(\HOLTyOp{num}\;\ensuremath{\rightarrow}\;\ensuremath{\alpha})\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}} is larger than that of $\alpha$, and hence we cannot derive $\phi$ is invariant for bisimulation for models with \HOLinline{(\HOLTyOp{num}\;\ensuremath{\rightarrow}\;\ensuremath{\alpha})\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}}-worlds from the fact that  $\phi$ is invariant for bisimulation for models with $\alpha$-worlds.
+If we could quantify over types (as we could in a theorem prover based on dependent type theory), then we can could define `invariant under bisimulation for models of every type', and hence prove the original statement of Van Benthem characterization theorem.
+
+For the proof of the two theorems above, the first one is immediate from \ref{2.47}, and the second one requires another critical lemma:
+
+\begin{thm}
+{\upshape\cite[Theorem 2.74, one direction]{Blackburn}}
+\label{2.74}
+\begin{holmath}
+  \ensuremath{\HOLTokenTurnstile}\HOLFreeVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{v}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{\ensuremath{\mathfrak{N}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{\ensuremath{\phi}}.\;\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLFreeVar{\ensuremath{\mathfrak{N}}},\HOLFreeVar{v}\ensuremath{\Vdash}\;\HOLBoundVar{\ensuremath{\phi}})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{U}\;\HOLBoundVar{J}.\\
+\;\;\;\;\;\;\;\;\;\;\;\HOLConst{ultrafilter}\;\HOLBoundVar{U}\;\HOLBoundVar{J}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\;\;\;\;\;\Pi\sb{\HOLBoundVar{U}}\;(\HOLTokenLambda{}\HOLBoundVar{j}.\;\HOLFreeVar{\ensuremath{\mathfrak{M}}})\HOLSymConst{,}\HOLTokenLeftbrace{}\HOLBoundVar{f}\;\HOLTokenBar{}\;(\HOLTokenLambda{}\HOLBoundVar{j}.\;\HOLFreeVar{w})\sim\sb{\HOLBoundVar{U}}\sp{\HOLConst{worlds}\;(\HOLTokenLambda{}\HOLBoundVar{j}.\;\HOLFreeVar{\ensuremath{\mathfrak{M}}})}\HOLBoundVar{f}\HOLTokenRightbrace{}\HOLSymConst{\ensuremath{\leftrightarroweq}}\Pi\sb{\HOLBoundVar{U}}\;(\HOLTokenLambda{}\HOLBoundVar{j}.\;\HOLFreeVar{\ensuremath{\mathfrak{N}}})\HOLSymConst{,}\HOLTokenLeftbrace{}\HOLBoundVar{g}\;\HOLTokenBar{}\\
+\;\;\;\;\;\;\;\;\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{j}.\;\HOLFreeVar{v})\sim\sb{\HOLBoundVar{U}}\sp{\HOLConst{worlds}\;(\HOLTokenLambda{}\HOLBoundVar{j}.\;\HOLFreeVar{\ensuremath{\mathfrak{N}}})}\HOLBoundVar{g}\HOLTokenRightbrace{}
+\end{holmath}
+\end{thm}
+The above lemma relies on \ref{2.73} and its mathematical proof can be found in \cite{Blackburn}.
 

### 2019-11-18 p-modalmodels/paper.tex
backup before revising

@@
 In HOL, we write \HOLinline{\HOLConst{FLT}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{\ensuremath{\Sigma}}} for the filtration of the model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} via the set \HOLinline{\HOLFreeVar{\ensuremath{\Sigma}}}.
-The formalization is almost a direct translation of the mathematical proof in \cite{Blackburn}.
+The formalization is almost a direct translation of the mathematical proof in Blackburn \emph{et al}~\cite{Blackburn}.
 For the only difference: the world set of the filtrated model using the standard proof is a set of equivalence classes of worlds.
@@
 \begin{holmath}
-  \ensuremath{\HOLTokenTurnstile}\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\sb{\mathrm{1}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\;\HOLBoundVar{w}\sb{\mathrm{2}}.\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLBoundVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}
+  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model}),(\HOLFreeVar{w}\sb{\mathrm{1}} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model})\;(\HOLBoundVar{w}\sb{\mathrm{2}} :\ensuremath{\beta}).\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLBoundVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}
 \end{holmath}
@@
 \;\;\;\;\;\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{j}.\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLBoundVar{y}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Dom})\;\HOLSymConst{\HOLTokenConj{}}\\
-\;\;\;\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{y}\;\HOLBoundVar{j}\;\HOLSymConst{=}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Fun}\;\HOLBoundVar{n}\;(\HOLConst{MAP}\;(\HOLTokenLambda{}\HOLBoundVar{\ensuremath{f\sb{U}}}.\;\HOLConst{CHOICE}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLBoundVar{j})\;\HOLBoundVar{zs})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U}\HOLTokenRightbrace{});\\
-\;\;\HOLFieldName{Pred}\;:=\;(\HOLTokenLambda{}\HOLBoundVar{p}\;\HOLBoundVar{zs}.\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Pred}\;\HOLBoundVar{p}\;(\HOLConst{MAP}\;(\HOLTokenLambda{}\HOLBoundVar{\ensuremath{f\sb{U}}}.\;\HOLConst{CHOICE}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLBoundVar{j})\;\HOLBoundVar{zs})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U})\HOLTokenRightrec{}
+\;\;\;\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{y}\;\HOLBoundVar{j}\;\HOLSymConst{=}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Fun}\;\HOLBoundVar{n}\;(\HOLConst{MAP}\;(\HOLTokenLambda{}\HOLBoundVar{\ensuremath{f\sb{U}}}.\;\HOLConst{CHOICE}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLBoundVar{j})\;\HOLBoundVar{zs})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\\
+\;\;\;\;\;\;\;\;\;\;\HOLFreeVar{U}\HOLTokenRightbrace{});\\
+\;\;\HOLFieldName{Pred}\;:=\\
+\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{p}\;\HOLBoundVar{zs}.\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Pred}\;\HOLBoundVar{p}\;(\HOLConst{MAP}\;(\HOLTokenLambda{}\HOLBoundVar{\ensuremath{f\sb{U}}}.\;\HOLConst{CHOICE}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLBoundVar{j})\;\HOLBoundVar{zs})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U})\HOLTokenRightrec{}
 \end{holmath}

### 2019-11-18 p-modalmodels/paper.tex
omit type parameter in p-modalmodels

@@
 \begin{holmath}
-  \ensuremath{\alpha}\;\HOLTyOp{form}\;=\;\HOLConst{VAR}\;\ensuremath{\alpha}\;\HOLTokenBar{}\;\HOLConst{DISJ}\;(\ensuremath{\alpha}\;\HOLTyOp{form})\;(\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLTokenBar{}\;\HOLSymConst{\ensuremath{\bot}}\;\HOLTokenBar{}\;(\HOLSymConst{\HOLTokenNeg{}})\;(\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLTokenBar{}\;\HOLConst{\ensuremath{\Diamond}}\;(\ensuremath{\alpha}\;\HOLTyOp{form})
+  \HOLTyOp{form}\;=\;\HOLConst{VAR}\;\HOLTyOp{num}\;\HOLTokenBar{}\;\HOLConst{DISJ}\;\HOLTyOp{form}\;\HOLTyOp{form}\;\HOLTokenBar{}\;\HOLSymConst{\ensuremath{\bot}}\;\HOLTokenBar{}\;(\HOLSymConst{\HOLTokenNeg{}})\;\HOLTyOp{form}\;\HOLTokenBar{}\;\HOLConst{\ensuremath{\Diamond}}\;\HOLTyOp{form}
 \end{holmath}
@@
   \ensuremath{\beta}\;\HOLTyOp{frame} = \HOLTokenLeftrec{}\;\HOLFieldName{world}\;{:}\;\ensuremath{\beta}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool};\;\HOLFieldName{rel}\;{:}\;\ensuremath{\beta}\;\ensuremath{\rightarrow}\;\ensuremath{\beta}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}\;\HOLTokenRightrec{}\\
-  (\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model} = \HOLTokenLeftrec{}\;\HOLFieldName{frame}\;{:}\;\ensuremath{\beta}\;\HOLTyOp{frame};\;\HOLFieldName{valt}\;{:}\;\ensuremath{\alpha}\;\ensuremath{\rightarrow}\;\ensuremath{\beta}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}\;\HOLTokenRightrec{}
+  \ensuremath{\beta}\;\HOLTyOp{model} = \HOLTokenLeftrec{}\;\HOLFieldName{frame}\;{:}\;\ensuremath{\beta}\;\HOLTyOp{frame};\;\HOLFieldName{valt}\;{:}\;\HOLTyOp{num}\;\ensuremath{\rightarrow}\;\ensuremath{\beta}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}\;\HOLTokenRightrec{}
 \end{holmath}
@@
 \begin{holmath}
-  (\HOLFreeVar{\ensuremath{\phi\sb{1}}} :\ensuremath{\alpha}\;\HOLTyOp{form})\equiv\sb{({:}\ensuremath{\beta})}(\HOLFreeVar{\ensuremath{\phi\sb{2}}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLTokenDefEquality{}\\
-\;\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model})\;(\HOLBoundVar{w} :\ensuremath{\beta}).\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{1}}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{2}}}\\[3mm]
+  (\HOLFreeVar{\ensuremath{\phi\sb{1}}} :\HOLTyOp{form})\equiv\sb{({:}\ensuremath{\beta})}(\HOLFreeVar{\ensuremath{\phi\sb{2}}} :\HOLTyOp{form})\;\HOLTokenDefEquality{}\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :\ensuremath{\beta}\;\HOLTyOp{model})\;(\HOLBoundVar{w} :\ensuremath{\beta}).\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{1}}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{2}}}\\[3mm]
 \end{holmath}
@@
 \begin{holmath}
-  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model}),(\HOLFreeVar{w} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
-\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :(\ensuremath{\alpha},\;\ensuremath{\beta}\;\HOLTyOp{list})\;\HOLTyOp{model})\;(\HOLBoundVar{r} :\ensuremath{\beta}\;\HOLTyOp{list}).\;\HOLConst{tree}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{frame}\;\HOLBoundVar{r}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{r}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}
+  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :\ensuremath{\beta}\;\HOLTyOp{model}),(\HOLFreeVar{w} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :\ensuremath{\beta}\;\HOLTyOp{list}\;\HOLTyOp{model})\;(\HOLBoundVar{r} :\ensuremath{\beta}\;\HOLTyOp{list}).\;\HOLConst{tree}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{frame}\;\HOLBoundVar{r}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{r}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}
 \end{holmath}
@@
 \begin{holmath}
-  \ensuremath{\HOLTokenTurnstile}\HOLConst{FINITE}\;(\HOLFreeVar{\ensuremath{\Phi}} :\ensuremath{\alpha}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{INFINITE}\;\ensuremath{{\cal{U}}}({:}\ensuremath{\beta})\;\HOLSymConst{\HOLTokenImp{}}\\
-\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{n} :\HOLTyOp{num}).\;\HOLConst{FINITE}\;{\HOLTokenLeftbrace{}\HOLBoundVar{\ensuremath{\phi}}\;\HOLTokenBar{}\;\HOLConst{DEG}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenLeq{}}\;\HOLBoundVar{n}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{prop_letters}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\Phi}}\HOLTokenRightbrace{}} /{\equiv\sb{(:\beta)}}
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{FINITE}\;(\HOLFreeVar{\ensuremath{\Phi}} :\HOLTyOp{num}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{INFINITE}\;\ensuremath{{\cal{U}}}({:}\ensuremath{\beta})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{n} :\HOLTyOp{num}).\\
+\;\;\;\;\;\;\;\;\;\;\;\HOLConst{FINITE}\;(\HOLTokenLeftbrace{}\HOLBoundVar{\ensuremath{\phi}}\;\HOLTokenBar{}\;\HOLConst{DEG}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenLeq{}}\;\HOLBoundVar{n}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{prop_letters}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\Phi}}\HOLTokenRightbrace{}\HOLSymConst{//E}\;(\HOLFreeVar{\ensuremath{\mu}} :\ensuremath{\beta}\;\HOLTyOp{itself}))
 \end{holmath}
@@
 \begin{holmath}
-  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model}),(\HOLFreeVar{w}\sb{\mathrm{1}} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
-\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :(\ensuremath{\alpha},\;\ensuremath{\beta}\;\HOLTyOp{list})\;\HOLTyOp{model})\;(\HOLBoundVar{v} :\ensuremath{\beta}\;\HOLTyOp{list}).\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{v}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}
+  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :\ensuremath{\beta}\;\HOLTyOp{model}),(\HOLFreeVar{w}\sb{\mathrm{1}} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :\ensuremath{\beta}\;\HOLTyOp{list}\;\HOLTyOp{model})\;(\HOLBoundVar{v} :\ensuremath{\beta}\;\HOLTyOp{list}).\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{v}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}
 \end{holmath}
@@
 \begin{holmath}
-  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model}),(\HOLFreeVar{w}\sb{\mathrm{1}} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
-\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model})\;(\HOLBoundVar{w}\sb{\mathrm{2}} :\ensuremath{\beta}).\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLBoundVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}
+  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :\ensuremath{\beta}\;\HOLTyOp{model}),(\HOLFreeVar{w}\sb{\mathrm{1}} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}} :\ensuremath{\beta}\;\HOLTyOp{model})\;(\HOLBoundVar{w}\sb{\mathrm{2}} :\ensuremath{\beta}).\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLBoundVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}
 \end{holmath}
@@
 \begin{holmath}
-  \HOLConst{ultraproduct_model}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLTokenDefEquality{}\\
+  \Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLTokenDefEquality{}\\
 \;\;\HOLTokenLeftrec{}\HOLFieldName{frame}\;:=\\
@@
 For our propose, we are only interested in adding the interpretation of finitely many nullary function symbols, which are also called \emph{constants}.
-We write \HOLinline{\HOLConst{is_expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}} to mean that \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}} is the result of adding each element in $A$ to \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} as a new constant.
+We write \HOLinline{\HOLFreeVar{is\HOLTokenUnderscore{}expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}} to mean that \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}} is the result of adding each element in $A$ to \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} as a new constant.
 Further, the function $f$ is a bijection between $\{0,\cdots,n-1\}$ and $A$, which is assumed to be finite, so that each nullary function symbol $c$ will be interpreted as \HOLinline{\HOLFreeVar{f}\;\HOLFreeVar{c}} in \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}}.
@@
 \begin{holmath}
-  \HOLConst{is_expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}\;\HOLTokenDefEquality{}\\
-\;\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}.\HOLFieldName{Dom}\;\HOLSymConst{=}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{BIJ}\;\HOLFreeVar{f}\;(\HOLConst{count}\;(\HOLConst{CARD}\;\HOLFreeVar{A}))\;\HOLFreeVar{A}\;\HOLSymConst{\HOLTokenConj{}}\\
-\;\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}.\HOLFieldName{Fun}\;\HOLSymConst{=}\\
-\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{c}\;\HOLBoundVar{l}.\;\HOLKeyword{if}\;\HOLBoundVar{c}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLConst{CARD}\;\HOLFreeVar{A}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{l}\;\HOLSymConst{=}\;[]\;\HOLKeyword{then}\;\HOLFreeVar{f}\;\HOLBoundVar{c}\;\HOLKeyword{else}\;\HOLConst{CHOICE}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom})\;\HOLSymConst{\HOLTokenConj{}}\\
-\;\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}.\HOLFieldName{Pred}\;\HOLSymConst{=}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Pred}
+  \HOLConst{expand}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{f}\;\HOLTokenDefEquality{}\\
+\;\;\HOLTokenLeftrec{}\HOLFieldName{Dom}\;:=\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom};\\
+\;\;\HOLFieldName{Fun}\;:=\\
+\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{c}\;\HOLBoundVar{l}.\;\HOLKeyword{if}\;\HOLBoundVar{c}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLConst{CARD}\;\HOLFreeVar{A}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{l}\;\HOLSymConst{=}\;[]\;\HOLKeyword{then}\;\HOLFreeVar{f}\;\HOLBoundVar{c}\;\HOLKeyword{else}\;\HOLConst{CHOICE}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom});\\
+\;\;\HOLFieldName{Pred}\;:=\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Pred}\HOLTokenRightrec{}
 \end{holmath}
 \end{defn}
-If \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}}, \HOLinline{\HOLFreeVar{A}} and \HOLinline{\HOLFreeVar{f}} are all fixed, then the model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}} such that \HOLinline{\HOLConst{is_expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}} is unique.
+If \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}}, \HOLinline{\HOLFreeVar{A}} and \HOLinline{\HOLFreeVar{f}} are all fixed, then the model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}} such that \HOLinline{\HOLFreeVar{is\HOLTokenUnderscore{}expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}} is unique.
 The only difference between a model and an expansion of it is the interpretation of function symbols.
@@
 %
-For every \HOLinline{\HOLFreeVar{A}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}}, with $|A|<n$ and for every $f:{\mathbb N}\to $ \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}}, there is a unique \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}} such that  \HOLinline{\HOLConst{is_expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}}.
+For every \HOLinline{\HOLFreeVar{A}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}}, with $|A|<n$ and for every $f:{\mathbb N}\to $ \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}}, there is a unique \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}'}}} such that  \HOLinline{\HOLFreeVar{is\HOLTokenUnderscore{}expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}}.
 %
@@
 \;\;\;\;\;\;\HOLConst{IMAGE}\;\HOLBoundVar{f}\;\ensuremath{{\cal{U}}}({:}\HOLTyOp{num})\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{A}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{CARD}\;\HOLBoundVar{A}\;\HOLSymConst{\HOLTokenLeq{}}\;\HOLFreeVar{n}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{A}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{Dom}\;\HOLSymConst{\HOLTokenConj{}}\\
-\;\;\;\;\;\;\HOLConst{is_expansion}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLBoundVar{A}\;\HOLBoundVar{\ensuremath{\mathfrak{M}'}}\;\HOLBoundVar{f}\;\HOLSymConst{\HOLTokenConj{}}\\
+\;\;\;\;\;\;\HOLConst{BIJ}\;\HOLBoundVar{f}\;(\HOLConst{count}\;(\HOLConst{CARD}\;\HOLBoundVar{A}))\;\HOLBoundVar{A}\;\HOLSymConst{\HOLTokenConj{}}\\
 \;\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{\ensuremath{\phi}}.\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\Gamma}}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{form_functions}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLTokenLeftbrace{}(\HOLBoundVar{c}\HOLSymConst{,}\HOLNumLit{0})\;\HOLTokenBar{}\;\HOLBoundVar{c}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLConst{CARD}\;\HOLBoundVar{A}\HOLTokenRightbrace{})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{ftype}\;\HOLBoundVar{x}\;\HOLBoundVar{\ensuremath{\Gamma}}\;\HOLSymConst{\HOLTokenConj{}}\\
-\;\;\;\;\;\;\HOLConst{consistent}\;\HOLBoundVar{\ensuremath{\mathfrak{M}'}}\;\HOLBoundVar{\ensuremath{\Gamma}}\;\HOLSymConst{\HOLTokenImp{}}\\
-\;\;\;\;\;\;\;\;\HOLConst{frealizes}\;\HOLBoundVar{\ensuremath{\mathfrak{M}'}}\;\HOLBoundVar{x}\;\HOLBoundVar{\ensuremath{\Gamma}}\\[3mm]
+\;\;\;\;\;\;\HOLConst{consistent}\;(\HOLConst{expand}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLBoundVar{A}\;\HOLBoundVar{f})\;\HOLBoundVar{\ensuremath{\Gamma}}\;\HOLSymConst{\HOLTokenImp{}}\\
+\;\;\;\;\;\;\;\;\HOLConst{frealizes}\;(\HOLConst{expand}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLBoundVar{A}\;\HOLBoundVar{f})\;\HOLBoundVar{x}\;\HOLBoundVar{\ensuremath{\Gamma}}\\[3mm]
 \end{holmath}
@@
   \HOLConst{shift_valuation}\;\HOLFreeVar{n}\;\HOLFreeVar{\ensuremath{\sigma}}\;\HOLFreeVar{f}\;\HOLTokenDefEquality{}\;(\HOLTokenLambda{}\HOLBoundVar{v}.\;\HOLKeyword{if}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLFreeVar{n}\;\HOLKeyword{then}\;\HOLFreeVar{f}\;\HOLBoundVar{v}\;\HOLKeyword{else}\;\HOLFreeVar{\ensuremath{\sigma}}\;(\HOLBoundVar{v}\;\HOLSymConst{\ensuremath{-}}\;\HOLFreeVar{n}))\\[2mm]
-  \ensuremath{\HOLTokenTurnstile}\HOLConst{is_expansion}\;(\HOLConst{mm2folm}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}})\;\HOLFreeVar{A}\;\HOLFreeVar{\ensuremath{\mathfrak{M}'}}\;\HOLFreeVar{f}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{valuation}\;(\HOLConst{mm2folm}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}})\;\HOLFreeVar{\ensuremath{\sigma}}\;\HOLSymConst{\HOLTokenConj{}}\\
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{BIJ}\;\HOLFreeVar{f}\;(\HOLConst{count}\;(\HOLConst{CARD}\;\HOLFreeVar{A}))\;\HOLFreeVar{A}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{valuation}\;(\HOLConst{mm2folm}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}})\;\HOLFreeVar{\ensuremath{\sigma}}\;\HOLSymConst{\HOLTokenConj{}}\\
 \;\;\;\;\;\HOLConst{form_functions}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLTokenLeftbrace{}(\HOLBoundVar{c}\sb{\mathrm{1}}\HOLSymConst{,}\HOLNumLit{0})\;\HOLTokenBar{}\;\HOLBoundVar{c}\sb{\mathrm{1}}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLConst{CARD}\;\HOLFreeVar{A}\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenImp{}}\\
-\;\;\;\;\;\;\;(\HOLFreeVar{\ensuremath{\mathfrak{M}'}},\HOLFreeVar{\ensuremath{\sigma}}\ensuremath{\vDash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenEquiv{}}\\
+\;\;\;\;\;\;\;(\HOLConst{expand}\;(\HOLConst{mm2folm}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}})\;\HOLFreeVar{A}\;\HOLFreeVar{f},\HOLFreeVar{\ensuremath{\sigma}}\ensuremath{\vDash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenEquiv{}}\\
 \;\;\;\;\;\;\;\;\;\;\HOLConst{mm2folm}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}},\HOLConst{shift_valuation}\;(\HOLConst{CARD}\;\HOLFreeVar{A})\;\HOLFreeVar{\ensuremath{\sigma}}\;\HOLFreeVar{f}\ensuremath{\vDash}\;\HOLConst{shift_form}\;(\HOLConst{CARD}\;\HOLFreeVar{A})\;\HOLFreeVar{\ensuremath{\phi}})
@@
 \;\;\HOLConst{FV}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLTokenLeftbrace{}\HOLFreeVar{x}\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\ensuremath{{\mathcal L}^1_\tau}}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\\
-\;\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :(\HOLTyOp{num},\;\ensuremath{\alpha})\;\HOLTyOp{model})\;(\HOLBoundVar{\ensuremath{\mathfrak{N}}} :(\HOLTyOp{num},\;\ensuremath{\beta})\;\HOLTyOp{model})\;(\HOLBoundVar{v} :\ensuremath{\beta})\;(\HOLBoundVar{w} :\ensuremath{\alpha}).\\
+\;\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :\ensuremath{\alpha}\;\HOLTyOp{model})\;(\HOLBoundVar{\ensuremath{\mathfrak{N}}} :\ensuremath{\beta}\;\HOLTyOp{model})\;(\HOLBoundVar{v} :\ensuremath{\beta})\;(\HOLBoundVar{w} :\ensuremath{\alpha}).\\
 \;\;\;\;\;\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{,}\HOLBoundVar{w}\HOLSymConst{\ensuremath{\leftrightarroweq}}\HOLBoundVar{\ensuremath{\mathfrak{N}}}\HOLSymConst{,}\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenImp{}}\\
@@
 \begin{holmath}
-  \ensuremath{\HOLTokenTurnstile}\HOLConst{FV}\;(\HOLFreeVar{\ensuremath{\delta}} :\HOLTyOp{folform})\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLTokenLeftbrace{}(\HOLFreeVar{x} :\HOLTyOp{num})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\ensuremath{{\mathcal L}^1_\tau}}\;\HOLFreeVar{\ensuremath{\delta}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\delta}}\;\;{}\sp{\sf f}{\equiv}\sb{({:}\ensuremath{\alpha})}\;\;{\sf ST}\sb{\HOLFreeVar{x}}\;(\HOLFreeVar{\ensuremath{\phi}} :\HOLTyOp{num}\;\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
+  \ensuremath{\HOLTokenTurnstile}\HOLConst{FV}\;(\HOLFreeVar{\ensuremath{\delta}} :\HOLTyOp{folform})\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLTokenLeftbrace{}(\HOLFreeVar{x} :\HOLTyOp{num})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\ensuremath{{\mathcal L}^1_\tau}}\;\HOLFreeVar{\ensuremath{\delta}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{\ensuremath{\delta}}\;\;{}\sp{\sf f}{\equiv}\sb{({:}\ensuremath{\alpha})}\;\;{\sf ST}\sb{\HOLFreeVar{x}}\;(\HOLFreeVar{\ensuremath{\phi}} :\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
 \;\;\;\;\;\;\;\HOLConst{invar4bisim}\;\HOLFreeVar{x}\;({:}\ensuremath{\alpha})\;({:}\ensuremath{\alpha})\;\HOLFreeVar{\ensuremath{\delta}} \\[3mm]
@@
 \;\;\;\;\;\;\;(\HOLFreeVar{\ensuremath{\delta}} :\HOLTyOp{folform})\;\HOLSymConst{\HOLTokenImp{}}\\
-\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\phi}} :\HOLTyOp{num}\;\HOLTyOp{form}).\;\HOLFreeVar{\ensuremath{\delta}}\;\;{}\sp{\sf f}{\equiv}\sb{({:}\ensuremath{\alpha})}\;\;{\sf ST}\sb{\HOLFreeVar{x}}\;\HOLBoundVar{\ensuremath{\phi}}
+\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\phi}} :\HOLTyOp{form}).\;\HOLFreeVar{\ensuremath{\delta}}\;\;{}\sp{\sf f}{\equiv}\sb{({:}\ensuremath{\alpha})}\;\;{\sf ST}\sb{\HOLFreeVar{x}}\;\HOLBoundVar{\ensuremath{\phi}}
 \end{holmath}

### 2020-03-09 p-modalmodels/response.md
Draft response.

We thank the reviewers for their careful reading of our paper.
We agree with the comments from Review 1 and 3 that we could profitably cite more work. In particular, apart from Harrison’s early work on mechanised model theory for predicate logic which we do mention, we might reasonably cite work on soundness and completeness for predicate logic (e.g., Blanchette, Popescu and Traytel’s nice recent work in Isabelle/HOL), as well as a sampling of modal logic mechanisations (such as de Wind’s thesis, a nice early example).

Naturally, we’d be happy to make Review 1’s other suggested changes (move our URL; change gothic M into something less intense).

Review 2 asks about how we might extend our work to multi-modal logics.  Extension to different flavours of unary diamonds would be straightforward: the syntax would feature indexed diamonds (◇_i), and models would include a corresponding indexed family of relations.
Review 2 also asks about including standard definitions of modal logics; our feeling is that the definitions in §2 (particularly nos. 1 to 3) are just those definitions. Though they are pretty-printed HOL, they link back to Blackburn et al’s originals, and we believe they are about as readable as in their textbook form.

Review 3 asks if we might solve our problem with type quantification by using a fixed base type.  Unfortunately, we don’t believe so.  Though we know that there are always finite models, we can’t carry out constructions like taking ultrafilters and keep a countable carrier set. We can’t then get the desired result.  We will try to make this point more clearly at the end of §7.

### 2020-03-10 p-modalmodels/response.md
add some sentences, maybe delete something later

@@
 Extension to different flavours of unary diamonds would be straightforward: the syntax would feature indexed diamonds (◇_i), and models would include a corresponding indexed family of relations.
+If we want to consider modal operators with any arity, we can change the constructor of formulas with modality to be ``DIAM n (form list)'', though it will later requires us to consider the well-formedness of formulas.
+(Maybe write less about this...)
+
 Review 2 also asks about including standard definitions of modal logics; our feeling is that the definitions in §2 (particularly nos.
@@
 Unfortunately, we don’t believe so.
-Though we know that there are always finite models, we can’t carry out constructions like taking ultrafilters and keep a countable carrier set.
-We can’t then get the desired result.
+Though we know that there are always finite models, it is not the fact that all the operations can be taken within a type.
+For instance, we can’t carry out constructions like taking ultrafilters and keep a countable carrier set.
+Similarly, we cannot take ultraproducts of models within the same type.
+Therefore, we can’t then get the desired result.
 We will try to make this point more clearly at the end of §7.

### 2020-03-11 p-modalmodels/response.md
More response edits

--- 2020-03-09 p-modalmodels/response.md
@@
 We agree with the comments from Review 1 and 3 that we could profitably cite more work.
-In particular, apart from Harrison’s early work on mechanised model theory for predicate logic which we do mention, we might reasonably cite work on soundness and completeness for predicate logic (e.g., Blanchette, Popescu and Traytel’s nice recent work in Isabelle/HOL), as well as a sampling of modal logic mechanisations (such as de Wind’s thesis, a nice early example).
+In particular, apart from Harrison’s early work on mechanised model theory for predicate logic which we do mention, we might reasonably cite work involving proof theory (e.g., Blanchette, Popescu and Traytel’s nice recent work in Isabelle/HOL), as well as a sampling of modal logic mechanisations (such as de Wind’s thesis, a nice early example).
 
@@
 Extension to different flavours of unary diamonds would be straightforward: the syntax would feature indexed diamonds (◇_i), and models would include a corresponding indexed family of relations.
+If we wanted to consider modal operators with any arity, we might change the diamond constructor take a list of formulas rather than just one.
+This would in turn require a well-formedness predicate to be defined over formulas to make sure that modalities were applied to the right number of arguments.
+With this syntactic nicety out of the way, we think the semantic arguments would lift fairly straightforwardly.
+
 Review 2 also asks about including standard definitions of modal logics; our feeling is that the definitions in §2 (particularly nos.
@@
 Unfortunately, we don’t believe so.
-Though we know that there are always finite models, we can’t carry out constructions like taking ultrafilters and keep a countable carrier set.
-We can’t then get the desired result.
+Though we know that there are always finite models, it is not the fact that all our desired operations can be taken within a type.
+For instance, we can’t carry out constructions like taking ultrafilters and keep a countable carrier set.
+Similarly, we cannot take ultraproducts of models and preserve cardinalities.
 We will try to make this point more clearly at the end of §7.
